### 2024-07-23 paper/paper.tex
added first rough draft of paper content

\documentclass{svproc}

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

% The following packages can be found on http:\\www.ctan.org
\usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{epsfig} % for postscript graphics files
\usepackage{mathptmx} % assumes new font selection scheme installed
\usepackage{times} % assumes new font selection scheme installed
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{caption}
\usepackage{subcaption}
\usepackage{array}
\usepackage{siunitx}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{makecell}

\title{\LARGE \bf
Laser Tracker Placement Optimization for highly flexible manufacturing systems
}

\author{Jan Baumgärtner$^{1}$,  Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
\thanks{*The authors would like to thank the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg for the financial support of the project within the InnovationsCampus Future Mobility.
}% <-this % stops a space
\thanks{$^{1}$Jan Baumgärtner, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
        Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
        {\tt\small jan.baumgaertner@kit.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\section{Formulation of the optimization problem}
To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
Similar to approaches such as \cite{ieee_sensors} we divide the visibility into two components.
Line of sight visibility and field of view visibility.
Both  are depicted in figure \ref{fig:visibility}.
\begin{figure}
        \centering
        \includegraphics[width=0.5\textwidth]{visibility.png}
        \caption{Visibility of a marker to a laser tracker}
        \label{fig:visibility}
\end{figure}
Mathematically we can define line of sight visibility (los visibility) as the following binary function:
\begin{equation}
    f(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
Note that $p_m$ is typically dependent on time as it is attached to a moving manufacuting system such as a robot.
The field of view objective can similarly be defined as a binary function:
\begin{equation}
    g(p_m,p_l) = \begin{cases}
    1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
However in practice it might make sense to not be close to the boundary of the field of view.
Instead one can define a continuous function that is 1 in the center of the field of view and 0 at the boundary.
The simplest way to encode this using the angle $\alpha$ between the normal of the marker and the line of sight of the laser tracker.
\begin{equation}
    g(p_m,p_l) = \cos(\alpha)
\end{equation}
The final objective function is then a weighted sum of the two components:
\begin{equation}
    h(p_m,p_l) = \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l)
    \label{eq:objective}
\end{equation}
where $\lambda$ is a weighting factor between 0 and 1 that can be adjusted to prioritize one component over the other.
After defining the objective function we can now define the design variables and constraints.
The design variables are the pose of the laser tracker $p_l$ and the pose of the marker $p_m$.
However the pose of these objects is not arbitrary.
For the purposes of this paper we assume that the marker is attached to the flange of the manufacturing system.
This means that we only define a Transformation matrix $T_{m}$ that describes the pose of the marker relative to the flange.
The final pose of the marker $p_m$ is then given by the flange transformation $T_{f}(q)$ where $q$ is the joint configuration of the robot as well as the transformation $T_{m}$:
\begin{equation}
    p_m = T_{f}(q)T_{m}
\end{equation}
While the matrix $T_m$ is a 4x4 matrix, we can parameterize it using the position of the marker $p_{p_m}$ and the orientation of the marker $p_{o_m}$.
This results in a 6 dimensional design space for the marker pose.
This space is further constraint to avoid any unrealistic marker positions (such as 10 meters above the robot).
Ideally the marer should be attached close to the surface of the manufacutring system to avoid any unneccessarily large constructions that might collide with the environment or cause vibrations.
This can be encoded using a cylindrical constraint on the position of the marker:
\begin{equation}
        \begin{split}
    \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
        0 \leq p_{p_m,z} &\leq h
        \end{split}
\end{equation}
where $r$ is the radius of the cylinder and $h$ is the height of the cylinder.
The marker of course also cant be within the mesh of the manufacturing system, but since this would result in a line of sight visibility of 0 we do not need to encode this constraint explicitly.

The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
This can be encoded using a simple constraint:
\begin{equation}
    h_{min} \leq p_{p_l,z} \leq h_{max}
\end{equation}
where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
The final design space is then 9 dimensional, 6 for the marker pose and 3 for the laser tracker pose.

\section{Optimization algorithm}
There are a large number of possible optimization algorithms that can be used to solve this problem.
Following the no free lunch theorem we want to pick the algorithm that makes the most use of the underlying structure of the problem. %TODO citation needed
In this case the problem is a continuous optimization problem with a relatively low number of design variables.
On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
These two properties make the problem well suited for the particle swarm optimization (pso) algorithm.
Particle swarm optimization (PSO) is a computational method for finding the optimal solution to a problem
 It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
 In PSO, a group of particles (also called agents or individu-als) move through a search space, according to the fol-lowing update rules:
\begin{equation}
        \begin{split}
        x_{i,j} &= x_{i,j} + v_{i,j} \\
    v_{i,j} &= wv_{i,j} + c_1r_1(p_{i,j}-x_{i,j}) + c_2r_2(p_{g,j}-x_{i,j})
        \end{split}
\end{equation}
where $x_{i,j}$ is the position of particle $i$ in dimension $j$, $v_{i,j}$ is the velocity of particle $i$ in dimension $j$, $w$ is the inertia weight,
 $c_1$ and $c_2$ are the cognitive and social coefficients, $r_1$ and $r_2$ are random numbers between 0 and 1, $p_{i,j}$ is the best position of particle $i$ in dimension $j$ and $p_{g,j}$
is the best position of the swarm in dimension $j$.
In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
The best particle is the one that maximizes the objective function defined in equation \ref{eq:objective}.
The hyperparameters we need to tune are collected in table \ref{tab:hyperparameters}.

\begin{table}  %TODO anpassen
        \centering
        \begin{tabular}{c|c}
                Hyperparameter & Value \\
                \hline
                Number of particles & 100 \\
                Number of iterations & 1000 \\
                Inertia weight & 0.9 \\
                Cognitive coefficient & 2.0 \\
                Social coefficient & 2.0 \\
        \end{tabular}
        \caption{Hyperparameters of the PSO algorithm}
        \label{tab:hyperparameters}
\end{table}

\section{Simulation Environment}
Evaluating the objective function requires a simulation of the manufacturing system.
The tricky part here is that since the manufacturing is actively manufacturing parts we not only need to simulate the behavior of the system itself but the changing shape of any parts that are being manufactured.
For this purpose we are using the pybullet-industrial package \cite{pybullet_industrial}.
This extension of the pybullet physics engine allows us to simulate not only the robot but also the manufacturing process.
For the purposes of this paper the markers are implemented as a pybullet industrial endeffector obect that is attached to the robot.

\section{Experiment Setup}

\end{document}

### 2024-07-23 paper/paper.tex
added code from ieee sensors paper

@@
         \centering
-        \includegraphics[width=0.5\textwidth]{visibility.png}
+        \includegraphics[width=0.5\textwidth]{figures/visibility.png}
         \caption{Visibility of a marker to a laser tracker}
@@
 where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
+The laser tracker also has a limited range of measurement.
+Here we can define a constraint that the distance between the laser tracker and the marker is within the range of the laser tracker:
+\begin{equation}
+    d_{min} \leq ||p_{p_m} - p_{p_l}|| \leq d_{max}
+\end{equation}
+where $d_{min}$ and $d_{max}$ are the minimum and maximum range of the laser tracker.
 The final design space is then 9 dimensional, 6 for the marker pose and 3 for the laser tracker pose.
@@
 The hyperparameters we need to tune are collected in table \ref{tab:hyperparameters}.
+Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
+Instead we can use a penalty function that penalizes any particles that violate the constraints.
+This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library. %TODO citation needed
+The penalty function can be defined as:
+\begin{equation}
+    p(x) =  \begin{cases}
+        0 & \text{if } x \text{ satisfies all constraints} \\
+        -\rho & \text{otherwise}
+    \end{cases}
+\end{equation}
+where $\rho$ is a large penalty factor.
 
@@
 For the purposes of this paper the markers are implemented as a pybullet industrial endeffector obect that is attached to the robot.
+A image of the simulation environment is shown in figure \ref{fig:simulation}.
+\begin{figure}
+        \centering
+        \includegraphics[width=0.5\textwidth]{figures/simulation.png}
+        \caption{Simulation environment}
+        \label{fig:simulation}
+\end{figure}
 

### 2024-07-23 code/multi_hist_comparison.py

@@ -1,0 +1,141 @@
+import matplotlib
+import numpy as np
+import pandas as pd
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+from paper_experiment import CustomEnv
+from paper_evaluation import DirectVisEnv, OrientationVisEnv, get_fov_from_intrinsic_matrix
+
+kit_deep_blue = "#144466"
+kit_red = "#B2372C"
+
+kit_green = "#009682"
+kit_yellow = "#EEB70D"
+
+light_blue = "#8EB1C7"
+pink = "#FFC6D9"
+purble = "#A288A6"
+shining_blue = "#70D6FF"
+
+colors = [kit_deep_blue, kit_green, kit_yellow,
+          light_blue, pink, purble, shining_blue]
+
+
+def load_camera_data(number_of_cameras):
+    optimal_states = np.load(
+        "optimal_states_"+str(number_of_cameras)+"_cameras.npy")
+    optimal_camera_states = np.split(
+        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
+
+    initial_states = np.load(
+        "initial_states_"+str(number_of_cameras)+"_cameras.npy")
+    initial_camera_states = np.split(
+        initial_states, len(initial_states)/6, axis=0)
+
+    focal_length = 0.1
+    intrinsic_matrix = [0.0]*9
+    intrinsic_matrix[0*3+0] = focal_length
+    intrinsic_matrix[1*3+1] = focal_length
+    intrinsic_matrix[2*3+2] = float(1)
+    intrinsic_matrix = tuple(intrinsic_matrix)
+
+    optimal_camera_configs = []
+    for camera_state in optimal_camera_states:
+        optimal_camera_configs.append([camera_state[:3], camera_state[3:6]])
+
+    initial_camera_configs = []
+    for i in range(len(initial_camera_states[0][0])):
+        initial_camera_configs.append([
+            [init_state[:3, i], init_state[3:6, i]] for init_state in initial_camera_states])
+
+    return optimal_camera_configs, initial_camera_configs, intrinsic_matrix
+
+
+def get_triangulation_data(number_of_cameras):
+    optimal_camera_configs, initial_camera_configs, intrinsic_matrix = load_camera_data(
+        number_of_cameras)
+
+    env = CustomEnv(False)
+    optimal_triangulation_error = env.configure_cameras([optimal_camera_configs], [[
+        intrinsic_matrix]*len(optimal_camera_configs)])
[…]

### 2024-07-23 code/optical_marker.py

@@ -1,0 +1,214 @@
+import os
+import numpy as np
+import pybullet as p
+import pybullet_industrial as pi
+
+import unittest
+
+
+def compute_angle_between_vectors(v1, v2):
+    """Computes the angle between two vectors
+
+    Args:
+        v1 (numpy.array): The first vector
+        v2 (numpy.array): The second vector
+
+    Returns:
+        float: The angle between the two vectors
+    """
+    v1_u = v1 / np.linalg.norm(v1)
+    v2_u = v2 / np.linalg.norm(v2)
+    return np.arccos(np.clip(np.dot(v1_u, v2_u), -1.0, 1.0))
+
+
+class OpticalMarker(pi.EndeffectorTool):
+    """Class which represents an optical marker
+    Args:
+        urdf_model (str): Path to the urdf model of the tool
+        start_position (list): Position of the tool in the world frame
+        start_orientation (list): Orientation of the tool in the world frame
+        marker_positions (list): Positions of the markers in the tool frame
+        coupled_robots (list): List of robots which are coupled to the tool
+        tcp_frame (list): Position of the tool center point in the tool frame
+        connector_frames (list): List of positions of the connectors in the tool frame
+    """
+
+    def __init__(self, urdf_model: str, start_position, start_orientation, marker_positions,
+                 coupled_robots=None, tcp_frame=None, connector_frames=None):
+        super().__init__(urdf_model, start_position, start_orientation,
+                         coupled_robots, tcp_frame, connector_frames)
+
+        self.marker_positions = marker_positions
+        steps = 50
+        particle_size = 0.03
+        width = particle_size*500
+        theta2 = np.linspace(-np.pi,  0, steps)
+        phi2 = np.linspace(0,  5 * 2*np.pi, steps)
+
+        for positions in marker_positions:
+            x_coord = particle_size * \
+                np.sin(theta2) * np.cos(phi2) + positions[0]
+            y_coord = particle_size * \
+                np.sin(theta2) * np.sin(phi2) + positions[1]
+            z_coord = particle_size * \
+                np.cos(theta2) + positions[2]
+
+            path = np.array([x_coord, y_coord, z_coord])
+            path_steps = len(path[0])
+            for i in range(1, path_steps):
+                current_point = path[:, i]
+                previous_point = path[:, i-1]
+                p.addUserDebugLine(current_point, previous_point,
+                                   lineColorRGB=[20/255, 68/255, 102/255],
+                                   lineWidth=width,
+                                   lifeTime=0,
[…]

### 2024-07-23 code/paper_evaluation.py

@@ -1,0 +1,150 @@
+from paper_experiment import CustomEnv, get_fov_from_intrinsic_matrix
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+kit_deep_blue = "#144466"
+kit_green = "#009682"
+kit_yellow = "#EEB70D"
+kit_red = "#B2372C"
+
+
+class DirectVisEnv(CustomEnv):
+
+    def _costs(self):
+        """Returns a cost matrix for each camera and marker.
+        """
+
+        # list of the visibility of each marker for each stereo camera system
+        visibility = self._compute_visibility(False)
+
+        camera_visibilities = [np.array(
+            [np.sum(vis) for vis in camera_visibility]) for camera_visibility in visibility]
+
+        return np.sum(camera_visibilities, axis=0)/(4*len(visibility[0][0]))
+
+
+class OrientationVisEnv(CustomEnv):
+
+    def _costs(self):
+        """Returns a cost matrix for each camera and marker.
+        """
+
+        # list of the visibility of each marker for each stereo camera system
+        visibility = self._compute_visibility(True, True)
+
+        camera_visibilities = [np.array(
+            np.max(np.abs(camera_visibility), axis=1)*np.pi) for camera_visibility in visibility]
+
+        return np.max(
+            np.max(np.abs(camera_visibilities), axis=0), axis=1)
+
+
+if __name__ == "__main__":
+    camera_radius = 2.5
+    number_of_cameras = 8
+    optimal_states = np.load(
+        "optimal_states_"+str(number_of_cameras)+"_cameras.npy")
+    optimal_camera_states = np.split(
+        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
+
+    initial_states = np.load(
+        "initial_states_"+str(number_of_cameras)+"_cameras.npy")
+    initial_camera_states = np.split(
+        initial_states, len(initial_states)/6, axis=0)
+
+    focal_length = 0.1
+    intrinsic_matrix = [0.0]*9
+    intrinsic_matrix[0*3+0] = focal_length
+    intrinsic_matrix[1*3+1] = focal_length
+    intrinsic_matrix[2*3+2] = float(1)
+    intrinsic_matrix = tuple(intrinsic_matrix)
+
+    optimal_camera_configs = []
+    for camera_state in optimal_camera_states:
+        optimal_camera_configs.append([camera_state[:3], camera_state[3:6]])
+
+    initial_camera_configs = []
+    for i in range(len(initial_camera_states[0][0])):
[…]

### 2024-07-23 code/paper_experiment.py

@@ -1,0 +1,542 @@
+import os
+
+import numpy as np
+import pybullet as p
+import pybullet_data
+import pybullet_industrial as pi
+
+from copy import deepcopy
+from optical_marker import OpticalMarker
+from particle_optimizer import particle_optimizer, particle_optimizer_log
+from rich.console import Console
+
+
+camera_radius = 2.5
+
+
+def quat_from_direction(direction: np.ndarray) -> np.ndarray:
+    """
+    Returns a quaternion that represents a coordinate system with its z-axis aligned with the given direction vector.
+
+    Args:
+        direction: A 3D direction vector.
+
+    Returns:
+        A quaternion representation of the desired coordinate system.
+    """
+    # Normalize the direction vector
+    direction = direction / np.linalg.norm(direction)
+
+    # If the direction is (0, 0, z), set the quaternion to (0, 0, 0, 1)
+    if direction[0] == 0 and direction[1] == 0:
+        return np.array([0, 0, 0, 1])
+
+    # Calculate the rotation axis and angle
+    rotation_axis = np.cross(np.array([0, 0, 1]), direction)
+    rotation_axis = rotation_axis / np.linalg.norm(rotation_axis)
+    rotation_angle = np.arccos(direction[2])
+
+    # Convert the rotation axis and angle to a quaternion
+    s = np.sin(rotation_angle / 2)
+    c = np.cos(rotation_angle / 2)
+    quaternion = np.array(
+        [s * rotation_axis[0], s * rotation_axis[1], s * rotation_axis[2], c])
+
+    return quaternion
+
+
+def align_with_center(particles_5d, path_center):
+    """Aligns 5d particles with the center of the path
+
+    Args:
+        particles_5d ([type]): The 5d particles
+        path_center ([type]): The center of the path
+
+    Returns:
+        [type]: The particles with their orientations aligned with the center of the path
+    """
+    center_aligned_particles = deepcopy(particles_5d)
+    number_of_cameras = int(len(particles_5d)/5)
+    for i in range(len(particles_5d[0])):
+        for j in range(number_of_cameras):
+
+            angles = particles_5d[j*5:j*5+2, i]
+            position = spherical_to_cartesian(
+                camera_radius, angles[0], angles[1])
+            direction = path_center-position
+
+            quaternion = quat_from_direction(direction)
+            center_aligned_particles[j*5+2:j*5+5,
[…]

### 2024-07-23 code/paper_setup_visualization.py

@@ -1,0 +1,26 @@
+from paper_experiment import CustomEnv
+import pybullet as p
+import pybullet_industrial as pi
+import numpy as np
+
+if __name__ == "__main__":
+
+    env = CustomEnv(True, 200)
+
+    focal_length = 0.1
+    intrinsic_matrix = [0.0]*9
+    intrinsic_matrix[0*3+0] = focal_length
+    intrinsic_matrix[1*3+1] = focal_length
+    intrinsic_matrix[2*3+2] = float(1)
+    intrinsic_matrix = tuple(intrinsic_matrix)
+    camera_state = np.array([0.0, 1, 0.6, np.pi/2, 0.0, 0.0])
+    camera_configs = []
+    pi.draw_coordinate_system(camera_state[:3],
+                              p.getQuaternionFromEuler(camera_state[3:6]))
+    camera_configs.append([camera_state[:3], camera_state[3:6]])
+
+    env.configure_cameras([camera_configs], [[
+                          intrinsic_matrix]*len(camera_configs)])
+
+    while True:
+        env.run_simulation()

### 2024-07-23 code/paper_visualization.py

@@ -1,0 +1,42 @@
+from paper_experiment import CustomEnv
+import pybullet as p
+import pybullet_industrial as pi
+import numpy as np
+
+if __name__ == "__main__":
+    camera_radius = 2.5
+    optimal_states = np.load("optimal_states_6_cameras.npy")
+    optimal_camera_states = np.split(
+        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
+
+    initial_states = np.load("initial_states_6_cameras.npy")
+    initial_camera_states = np.split(
+        initial_states, len(initial_states)/6, axis=0)
+    env = CustomEnv(True)
+
+    focal_length = 0.1
+    intrinsic_matrix = [0.0]*9
+    intrinsic_matrix[0*3+0] = focal_length
+    intrinsic_matrix[1*3+1] = focal_length
+    intrinsic_matrix[2*3+2] = float(1)
+    intrinsic_matrix = tuple(intrinsic_matrix)
+
+    camera_configs = []
+    for camera_state in optimal_camera_states:
+        pi.draw_coordinate_system(camera_state[:3],
+                                  p.getQuaternionFromEuler(camera_state[3:6]))
+        camera_configs.append([camera_state[:3], camera_state[3:6]])
+
+    env.configure_cameras([camera_configs], [[
+                          intrinsic_matrix]*len(camera_configs)])
+
+    for camera_state in initial_camera_states:
+        for i in range(len(camera_state[0])):
+            pi.draw_coordinate_system(camera_state[:3, i],
+                                      p.getQuaternionFromEuler(
+                                          camera_state[3:6, i]),
+                                      length=0.05, width=1.0)
+    env._compute_visibility()
+    env.run_simulation()
+    while True:
+        p.stepSimulation()

### 2024-07-23 code/particle_optimizer.py

@@ -1,0 +1,123 @@
+"""
+This file contains the implementation of a simple particle swarm optimizer.
+The algorithm is based on the paper "Particle Swarm Optimization"
+by James Kennedy and Russell Eberhart.
+"""
+
+import numpy as np
+from copy import deepcopy
+
+
+def particle_optimizer(objective, start_positions, start_velocities, max_iter):
+    """Simple particle swarm optimization implementation.
+    While the algorithm itself is unconstrained. Constrained optimization can be performed by
+    including a constraint penalty a*max(0,g(x)) into the objective function.
+
+    Args:
+        objective ([type]): A objective function to be minimized.
+                            The function has to be vectorized such that the objective of all
+                            states can be computed at once.
+        start_positions ([type]): The initial state of the particles with dimension
+                                  (number_of_particle_states,number_of_particles)
+        start_velocities ([type]): The initial velocity of the particles with the same
+                                   dimension as the start_positions
+        max_iter ([type]): the number of iterations before the system terminates
+
+    Returns:
+        [type]: The current global optimum state
+        float: the objective function evaluated at the current global optimum state
+    """
+    c1 = c2 = 0.1
+    w = 0.8
+    X = deepcopy(start_positions)
+    V = deepcopy(start_velocities)
+
+    pbest = deepcopy(start_positions)
+    pbest_obj = objective(X)
+    gbest = pbest[:, pbest_obj.argmin()]
+    gbest_obj = pbest_obj.min()
+
+    for i in range(max_iter):
+        print("iteration", i)
+        r1, r2 = np.random.rand(2)
+        V = w * V + c1*r1*(pbest - X) + c2*r2*(gbest.reshape(-1, 1)-X)
+        X = X + V
+        obj = objective(X)
+        pbest[:, (pbest_obj >= obj)] = X[:, (pbest_obj >= obj)]
+        pbest_obj = np.array([pbest_obj, obj]).min(axis=0)
+        gbest = pbest[:, pbest_obj.argmin()]
+        gbest_obj = pbest_obj.min()
+
+    return gbest, gbest_obj
+
+
+def particle_optimizer_log(objective, start_positions, start_velocities, max_iter):
+    """Simple particle swarm optimization implementation. This variate retruns the history of
+       the optimization process.
+
+    Args:
+        objective ([type]): A objective function to be minimized.
[…]

### 2024-07-23 code/singular_orientation_exp.py

@@ -1,0 +1,41 @@
+import numpy as np
+from paper_experiment import CustomEnv
+
+
+if __name__ == "__main__":
+    env = CustomEnv(False)
+
+    focal_length = 0.1
+    intrinsic_matrix = [0.0]*9
+    intrinsic_matrix[0*3+0] = focal_length
+    intrinsic_matrix[1*3+1] = focal_length
+    intrinsic_matrix[2*3+2] = float(1)
+    intrinsic_matrix = tuple(intrinsic_matrix)
+
+    base_camera_state = [[0, 0, 0], [0, 0, 0]]
+
+    rotating_camera_state = [[1, 0, 0], [0, 0, 0]]
+
+    steps = 100
+
+    # create an array of angles from -pi/2 to pi/2  in n steps
+    camera_angles = np.linspace(-np.pi/2, np.pi/2, steps)
+
+    singular_values = []
+
+    for camera_angle in camera_angles:
+        rotating_camera_state[1][1] = camera_angle
+
+        singular_value = env.configure_cameras([[base_camera_state, rotating_camera_state]], [[
+            intrinsic_matrix, intrinsic_matrix]])
+
+        singular_values.append(singular_value)
+
+    env.close()
+
+    import matplotlib.pyplot as plt
+    plt.plot(camera_angles, singular_values)
+    plt.title("Singular values of the camera configurations")
+    plt.xlabel("Camera angle relative to base camera [rad]")
+    plt.ylabel("Singular values")
+    plt.show()

### 2024-07-23 code/test_camera_model.py

@@ -1,0 +1,63 @@
+import os
+
+import numpy as np
+import pybullet as p
+import pybullet_industrial as pi
+
+
+if __name__ == "__main__":
+    # setup a pybullet simulation
+    p.connect(p.GUI)
+
+    # pybullet load marker.urdf
+    dirname = os.path.dirname(__file__)
+    urdf_file = os.path.join(dirname, 'marker.urdf')
+    # load urdf
+    p.loadURDF(urdf_file, [0.0, 0, 0], [0, 0, 0, 1], useFixedBase=True)
+
+    camera_parameters = {}
+    fov = 60
+    img_width = 640
+    img_height = 480
+    near_plane = 0.1
+    far_plane = 100
+    aspect_ratio = 1
+    focal_length = 10
+
+    projection_matrix = p.computeProjectionMatrixFOV(
+        fov, aspect_ratio, near_plane, far_plane)
+
+    first_camera_pos = np.array([-0., 0, -1])
+    second_camera_pos = np.array([-0., 0., -1])
+
+    first_camera_rot = [0, 0, 0, 1]
+    second_camera_rot = p.getQuaternionFromEuler([0, 0, 0])
+
+    pi.draw_coordinate_system(first_camera_pos, first_camera_rot)
+    pi.draw_coordinate_system(second_camera_pos, second_camera_rot)
+
+    # convert quaternion to rotation matrix
+    first_camera_rot_mat = p.getMatrixFromQuaternion(first_camera_rot)
+    first_camera_rot_mat = np.array(first_camera_rot_mat).reshape(3, 3)
+    second_camera_rot_mat = p.getMatrixFromQuaternion(second_camera_rot)
+    second_camera_rot_mat = np.array(second_camera_rot_mat).reshape(3, 3)
+
+    # calculate camera target position in world frame
+    first_camera_target_pos = first_camera_pos+focal_length * \
+        first_camera_rot_mat.dot(np.array([0, 0, 1]))
+    second_camera_target_pos = second_camera_pos+focal_length * \
+        second_camera_rot_mat.dot(np.array([0, 0, 1]))
+
+    # Setup the view matrices for both camera
+    first_camera_view_matrix = p.computeViewMatrix(
+        first_camera_pos, first_camera_target_pos, np.array([0, 1, 0]))
+    second_camera_view_matrix = p.computeViewMatrix(
+        second_camera_pos, second_camera_target_pos, np.array([0, 1, 0]))
+
+    first_image = p.getCameraImage(
[…]

### 2024-07-23 code/transformer_cell_simulation-main/Repositioning_Test.py
added sfb sim files for  experiment setup

@@ -1,0 +1,89 @@
+import math
+import os
+import time
+import numpy as np
+import pybullet as p
+import pybullet_data
+import pybullet_industrial as pi
+from pybullet_industrial import GCodeProcessor
+def execute_gcode(processor, command):
+    processor.gcode = processor.read_gcode(command)
+    for _ in iter(processor):
+        # Execute the simulation steps
+
+        for _ in range(50):
+            p.stepSimulation()
+            time.sleep(0.01)
+if __name__ == '__main__':
+    dirname = os.path.dirname(__file__)
+    comau_urdf = os.path.join(dirname,
+                              'robot_descriptions', 'comau_nj290_robotNC.urdf')
+    plate_urdf = os.path.join(dirname,
+                                'robot_descriptions', 'plate.urdf')
+    plane_urdf = os.path.join(dirname,
+                                 'objects', 'plane.urdf')
+
+    physics_client = p.connect(p.GUI, options=  '--width=1920 --height=1060 --background_color_red=0.8 --background_color_green=0.8 --background_color_blue=0.83')
+    p.configureDebugVisualizer(p.COV_ENABLE_GUI, 0) # 0: no gui, 1, basic gui
+
+    #p.resetDebugVisualizerCamera(cameraDistance=6.6, cameraYaw=-7, cameraPitch=-35,
+    #                             cameraTargetPosition=np.array([-0.5, 0, -1]))
+
+    p.setPhysicsEngineParameter(numSolverIterations=10000)
+
+    p.setAdditionalSearchPath(pybullet_data.getDataPath())
+
+    p.loadURDF(plane_urdf, [0, 0, 0], p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True)
+    #load Robots__________________________________________________
+    start_pos = np.array([-3.6353162, -0.6, 0])
+    start_orientation = p.getQuaternionFromEuler([0, 0, 0])
+    robot = pi.RobotBase(comau_urdf, start_pos, start_orientation)
+    start_pos2 = np.array([-0.12, -2.59, 0])
+    start_orientation2 = p.getQuaternionFromEuler([0, 0, -np.pi/2])
+    robot2 = pi.RobotBase(comau_urdf, start_pos2, start_orientation2)
+    plate = pi.EndeffectorTool(plate_urdf, np.array([-0.12,-4,2]),
[…]

### 2024-07-23 code/transformer_cell_simulation-main/main.py

@@ -1,0 +1,71 @@
+import math
+import os
+import time
+import numpy as np
+import pybullet as p
+import pybullet_data
+import pybullet_industrial as pi
+
+
+
+if __name__ == '__main__':
+    screwing = True
+    placing = True
+
+    dirname = os.path.dirname(__file__)
+    comau_urdf = os.path.join(dirname,
+                              'robot_descriptions', 'comau_nj290_robot.urdf')
+    gripper_urdf = os.path.join(dirname,
+                              'robot_descriptions', 'gripper_cad.urdf')
+    screwDriver_urdf = os.path.join(dirname,
+                              'robot_descriptions', 'screwDriver.urdf')
+    plane_urdf = os.path.join(dirname,
+                                 'objects', 'plane.urdf')
+
+    physics_client = p.connect(p.GUI, options=  '--width=1920 --height=1060 --background_color_red=1 --background_color_green=1 --background_color_blue=1')
+    p.configureDebugVisualizer(p.COV_ENABLE_GUI, 0) # 0: no gui, 1, basic gui
+
+    p.resetDebugVisualizerCamera(cameraDistance=6.6, cameraYaw=-7, cameraPitch=-35,
+                                 cameraTargetPosition=np.array([-0.5, 0, -1]))
+
+    p.setPhysicsEngineParameter(numSolverIterations=10000)
+
+    p.setAdditionalSearchPath(pybullet_data.getDataPath())
+
+
+    p.loadURDF(plane_urdf, [0, 0, 0], p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True, globalScaling=0.22)
+    #load Robots__________________________________________________
+    start_pos = np.array([-3.6353162, -0.6, 0])
+    start_orientation = p.getQuaternionFromEuler([0, 0, 0])
+    robot = pi.RobotBase(comau_urdf, start_pos, start_orientation)
+    start_pos2 = np.array([-0.12, -2.59, 0])
+    start_orientation2 = p.getQuaternionFromEuler([0, 0, np.pi])
+    robot2 = pi.RobotBase(comau_urdf, start_pos2, start_orientation2)
+    robot.set_joint_position(
+        {'q2': np.deg2rad(-15.0), 'q3': np.deg2rad(-90.0)})
+    robot2.set_joint_position(
+        {'q2': np.deg2rad(-15.0), 'q3': np.deg2rad(-90.0)})
+
[…]

### 2024-07-23 code/transformer_cell_simulation-main/mainC02.py

@@ -1,0 +1,249 @@
+import math
+import os
+import time
+import numpy as np
+import pybullet as p
+import pybullet_data
+import pybullet_industrial as pi
+
+def linear_interpolation_orn(start_point: np.array, end_point: np.array,
+                             start_orn: np.array, end_orn: np.array, samples: int):
+    """Performs a linear interpolation betwenn two points in 3D space
+
+    Args:
+        start_point (np.array): The start point of the interpolation
+        end_point (np.array): The end point of the interpolation
+        samples (int): The number of samples used to interpolate
+
+    Returns:
+        ToolPath: A ToolPath object of the interpolated path
+    """
+    final_path = np.linspace(start_point, end_point, num=samples)
+    final_orns = np.linspace(start_orn, end_orn, num=samples)
+    return pi.ToolPath(positions=final_path.transpose(), orientations= final_orns.transpose())
+
+def move_along_path(endeffector: pi.EndeffectorTool, path: pi.ToolPath, stop=True):
+    """Moving a designated endeffector along the provided path.
+    Args:
+        endeffector (pi.EndeffectorTool): Endeffector to be moved.
+        path (pi.ToolPath): Array of points defining the path.
+        stop (bool, optional): Whether or not to stop at the end of the movement.
+    """
+    #path.draw()
+    for positions, orientations, tool_path in path:
+        endeffector.set_tool_pose(positions, orientations)
+        for _ in range(20):
+            p.stepSimulation()
+    if stop:
+        for _ in range(50):
+            p.stepSimulation()
+def move_lin(endeffector: pi.EndeffectorTool, target_pos, target_orn=None, step_size=0.1):
+    pos, orn = endeffector.get_tool_pose("tcp")
+    if target_orn == None:
+        target_orn = orn
+
+    steps=round(np.linalg.norm(target_pos - pos)/step_size)
+    path=linear_interpolation_orn(pos, target_pos,
+                                       orn,
+                                       target_orn,
+                                       steps)
[…]

### 2024-07-23 code/paper_experiment.py
updated simulation

@@ -438,9 +438,9 @@ if __name__ == "__main__":
 
     # set up first environment with only orientation
 
-    number_of_cameras = 6
 
-    def run_experiment(number_of_cameras):
+
+    def run_experiment():
 
         def objective_function(particles):
 
@@ -494,48 +494,21 @@ if __name__ == "__main__":
 
         n_particles = 100
         initial_population = np.random.rand(
-            2, n_particles*number_of_cameras)*0.3*np.pi
+            9, n_particles)*0.3*np.pi
 
-        initial_5d_population = get_5d_from_2d(
-            initial_population.reshape(2*number_of_cameras, n_particles))
 
-        paths = env.paths
-        path_center = np.mean([np.mean(path.positions, axis=1)
-                               for path in paths], axis=0)
-        initial_5d_population = align_with_center(
-            initial_5d_population, path_center)
-        np.save("initial_states_"+str(number_of_cameras)+"_cameras.npy", spherical_particles_to_cartesian_particles(
-            initial_5d_population, camera_radius))
+        np.save("initial_states.npy", spherical_particles_to_cartesian_particles(
+            initial_population, camera_radius))
 
         start_velocities = np.random.randn(
-            2, number_of_cameras*n_particles) * 0.5
+            9, n_particles) * 0.5
 
         with console.status("Finding optimal camera placements"):
             particle_log, obj_log = particle_optimizer_log(no_orientation_spherical_stereo_objective_function, initial_population,
                                                            start_velocities, max_iter=100)
         env.close()
 
-        # set up second environment with orientation
-        env = CustomEnv()
 
-        grouped_particle_log = particle_log[-1].reshape(
-            2*number_of_cameras, n_particles)
-        initial_population = get_5d_from_2d(grouped_particle_log)
-
-        # orient initial population towards the center
-        initial_population = align_with_center(initial_population, path_center)
-
-        initial_velocities = np.random.randn(
-            5*number_of_cameras, n_particles) * 0.5
-
-        with console.status("Finding optimal camera placements"):
-            optimal_states = particle_optimizer(spherical_stereo_objective_function, initial_population,
-                                                initial_velocities, max_iter=200)
-
-        # save results
-        np.save("optimal_states_"+str(number_of_cameras)+"_cameras.npy", spherical_particles_to_cartesian_particles(
-            optimal_states[0].reshape((len(optimal_states[0]), 1)), camera_radius))
-        env.close()
[…]

### 2024-07-23 code/paper_experiment.py
renamed transformer cell and started integrating into marker code

@@ -214,16 +214,32 @@ class CustomEnv:
         return visibility
 
     def _populate_cell(self):
-        dirname = os.path.dirname(__file__)
-        urdf_file = os.path.join(dirname, 'ground.urdf')
-        p.loadURDF(urdf_file, [0, 0, 0], useFixedBase=True)
+        dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects')
+        #load Objects__________________________________________________
+        plane_urdf = os.path.join(dirname,'plane.urdf')
+        p.loadURDF(plane_urdf, [0, 0, 0], p.getQuaternionFromEuler([0, 0, 0]),
+                   useFixedBase=True, globalScaling=0.22)
 
-        cylinder_id = p.createCollisionShape(
-            p.GEOM_CYLINDER, radius=0.5, height=1)
-        cylinder_color = p.createVisualShape(
-            p.GEOM_CYLINDER, radius=0.5, length=1, rgbaColor=[196/255, 208/255, 207/255, 1], specularColor=[0, 0, 0])
-        p.createMultiBody(0, cylinder_id, cylinder_color,
-                          self.cylinder_center, p.getQuaternionFromEuler([np.pi/2, 0, 0]))
+
+        fofa_path = os.path.join(dirname, 'InRePro', 'InRePro.urdf')
+        p.loadURDF(fofa_path, [0,0,0], p.getQuaternionFromEuler([np.pi/2,0,0]),
+                   useFixedBase=True, globalScaling=0.001)
+
+        ponticon_path = os.path.join(dirname,'ponticon', 'ponticon.urdf')
+        p.loadURDF(ponticon_path, [0,0,0], p.getQuaternionFromEuler([np.pi/2,0,0]),
+                   useFixedBase=True, globalScaling=0.001)
+
+        dmg_path = os.path.join(dirname,'DMG', 'DMG.urdf')
+        p.loadURDF(dmg_path, [0,0,0], p.getQuaternionFromEuler([np.pi/2,0,0]),
+                   useFixedBase=True, globalScaling=0.001)
+
+        mm_path = os.path.join(dirname,  'modules', 'milling_module.urdf')
+        p.loadURDF(mm_path, [-0.6,-4.6,0.95], p.getQuaternionFromEuler([0,0,np.pi/2]),
+                   useFixedBase=True, globalScaling=0.001)
+
+        table_path = os.path.join(dirname,  'modules', 'table.urdf')
[…]

### 2024-07-23 paper/paper.tex
added sfb sim files for  experiment setup
added visibility plot to paper
started experiment section

@@
         \centering
-        \includegraphics[width=0.5\textwidth]{figures/visibility.png}
-        \caption{Visibility of a marker to a laser tracker}
+        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
+        \caption{Field of view (VOW) angle range and line of sight (LOS) components of marker visibility}
         \label{fig:visibility}
@@
         \centering
-        \includegraphics[width=0.5\textwidth]{figures/simulation.png}
-        \caption{Simulation environment}
+        \includegraphics[width=\textwidth]{figures/simulation.png}
+        \caption{Simulation environment used for the optimization}
         \label{fig:simulation}
@@
 \section{Experiment Setup}
+To test the performance of our optimiser, we need to define a set of test cases.
+These consist of different trajectories of the two robots shown in figure \ref{fig:simulation}.
+Both robots are Comau NJ290 industrial robots with a reach of 2.9 metres.
+The right robot is used to service the CNC machine and the additive manufacturing system, while the left robot is used to perform a high-precision manufacturing task.
+To properly control this task, the left robot needs to be tracked by the laser tracker.
+We test whether our algorithm is able to find a placement of the laser tracking system and markers that can track the robot over its entire trajectory.
+We compare this to the performance of the original population of marker and laser tracker placements used to start the optimisation.
 

### 2024-07-23 code/paper_experiment.py
started adding markers

@@ -262,10 +262,24 @@ class CustomEnv:
         robot2 = pi.RobotBase(comau_urdf, start_pos2, start_orientation2)
 
 
+        marker = OpticalMarker(os.path.join(dirname, 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1],
+                                 [[0.1, 0.1, 0.1],
+                                  [0.1, -0.1, 0.1],
+                                  [-0.1, 0.1, 0.1],
+                                  [-0.1, -0.1, 0.1]])
+        marker.couple
+
+        second_endeffector = OpticalMarker(os.path.join(dirname, 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1],
+                                    [[0.1, 0.1, 0.1],
+                                    [0.1, -0.1, 0.1],
+                                    [-0.1, 0.1, 0.1],
+                                    [-0.1, -0.1, 0.1]])
+
+        second_endeffector.couple(robot2, [0, 0, 0], [0, 0, 0, 1])
 
         # TODO define end effectors, and paths to return
 
-        return [marker, second_marker], [upward, downward]
+        return [marker,second_endeffector], [upward]
 
     def reset(self):
         """Resets the environment to the initial state

### 2024-07-24 code/paper_experiment.py
started with new cleaner environment class version

@@ -158,9 +158,8 @@ def get_fov_from_intrinsic_matrix(intrinsic_matrix):
     return 2*np.arctan(0.5/intrinsic_matrix[0])
 
 
-class CustomEnv:
-    radius = 0.6
-    cylinder_center = np.array([-1, 0, 0.5])
+
+class LaserTrackerEnv:
 
     def __init__(self,  rendering=False, steps=20):
         if rendering is False:
@@ -176,44 +175,7 @@ class CustomEnv:
         self._time_step_length = 0.1
         p.setTimeStep(self._time_step_length)
 
-        self.number_of_steps = steps
-        self.markers, self.paths = self._setup_cell()
-        self._populate_cell()
 
-        self.stereo_camera_states = []
-        self.intrinsic_matrices = []
-        self.min_singular_value = []
-
-        self._populate_cell()
-        p.configureDebugVisualizer(p.COV_ENABLE_RENDERING, 1)
-
-    def close(self):
-        """disconnects the environment from the physics server
-        """
-        p.disconnect(self.physics_client)
-
-    def _compute_visibility(self, orientation=True, orientation_only=False):
-        """ Function that computes the visibility of each marker for each stereo camera set."""
-        flattened_states = [
-            item for sublist in self.stereo_camera_states for item in sublist]
-        quaternion_states = [[state[0], p.getQuaternionFromEuler(
-            state[1])] for state in flattened_states]
-
-        if orientation:
-            fovs = [get_fov_from_intrinsic_matrix(intrinsic_matrix)
-                    for sublist in self.intrinsic_matrices for intrinsic_matrix in sublist]
-        else:
-            fovs = None
-
-        visibility = []
-        for marker in self.markers:
-            flattened_visibility = marker.compute_visibility(
-                quaternion_states, fovs, orientation_only)
-            visibility.append(
-                reshape(self.stereo_camera_states, flattened_visibility))
-        return visibility
-
-    def _populate_cell(self):
         dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects')
         #load Objects__________________________________________________
         plane_urdf = os.path.join(dirname,'plane.urdf')
@@ -241,45 +203,31 @@ class CustomEnv:
         p.loadURDF(table_path, [-0.6, -4.6, 0.95], p.getQuaternionFromEuler([0, 0, np.pi/2]),
                    useFixedBase=True, globalScaling=0.001)
 
-    def _setup_cell(self):
[…]

### 2024-07-24 code/paper_evaluation.py
cleaned up repo

@@ -1,150 +1,0 @@
-from paper_experiment import CustomEnv, get_fov_from_intrinsic_matrix
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-kit_deep_blue = "#144466"
-kit_green = "#009682"
-kit_yellow = "#EEB70D"
-kit_red = "#B2372C"
-
-
-class DirectVisEnv(CustomEnv):
-
-    def _costs(self):
-        """Returns a cost matrix for each camera and marker.
-        """
-
-        # list of the visibility of each marker for each stereo camera system
-        visibility = self._compute_visibility(False)
-
-        camera_visibilities = [np.array(
-            [np.sum(vis) for vis in camera_visibility]) for camera_visibility in visibility]
-
-        return np.sum(camera_visibilities, axis=0)/(4*len(visibility[0][0]))
-
-
-class OrientationVisEnv(CustomEnv):
-
-    def _costs(self):
-        """Returns a cost matrix for each camera and marker.
-        """
-
-        # list of the visibility of each marker for each stereo camera system
-        visibility = self._compute_visibility(True, True)
-
-        camera_visibilities = [np.array(
-            np.max(np.abs(camera_visibility), axis=1)*np.pi) for camera_visibility in visibility]
-
-        return np.max(
-            np.max(np.abs(camera_visibilities), axis=0), axis=1)
-
-
-if __name__ == "__main__":
-    camera_radius = 2.5
-    number_of_cameras = 8
-    optimal_states = np.load(
-        "optimal_states_"+str(number_of_cameras)+"_cameras.npy")
-    optimal_camera_states = np.split(
-        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
-
-    initial_states = np.load(
-        "initial_states_"+str(number_of_cameras)+"_cameras.npy")
-    initial_camera_states = np.split(
-        initial_states, len(initial_states)/6, axis=0)
-
-    focal_length = 0.1
-    intrinsic_matrix = [0.0]*9
-    intrinsic_matrix[0*3+0] = focal_length
-    intrinsic_matrix[1*3+1] = focal_length
-    intrinsic_matrix[2*3+2] = float(1)
-    intrinsic_matrix = tuple(intrinsic_matrix)
-
-    optimal_camera_configs = []
-    for camera_state in optimal_camera_states:
-        optimal_camera_configs.append([camera_state[:3], camera_state[3:6]])
-
-    initial_camera_configs = []
-    for i in range(len(initial_camera_states[0][0])):
[…]

### 2024-07-24 code/paper_experiment.py

@@ -2,160 +2,14 @@ import os
 
 import numpy as np
 import pybullet as p
-import pybullet_data
 import pybullet_industrial as pi
 
-from copy import deepcopy
+
 from optical_marker import OpticalMarker
 from particle_optimizer import particle_optimizer, particle_optimizer_log
-from rich.console import Console
 
 
-camera_radius = 2.5
 
-
-def quat_from_direction(direction: np.ndarray) -> np.ndarray:
-    """
-    Returns a quaternion that represents a coordinate system with its z-axis aligned with the given direction vector.
-
-    Args:
-        direction: A 3D direction vector.
-
-    Returns:
-        A quaternion representation of the desired coordinate system.
-    """
-    # Normalize the direction vector
-    direction = direction / np.linalg.norm(direction)
-
-    # If the direction is (0, 0, z), set the quaternion to (0, 0, 0, 1)
-    if direction[0] == 0 and direction[1] == 0:
-        return np.array([0, 0, 0, 1])
-
-    # Calculate the rotation axis and angle
-    rotation_axis = np.cross(np.array([0, 0, 1]), direction)
-    rotation_axis = rotation_axis / np.linalg.norm(rotation_axis)
-    rotation_angle = np.arccos(direction[2])
-
-    # Convert the rotation axis and angle to a quaternion
-    s = np.sin(rotation_angle / 2)
-    c = np.cos(rotation_angle / 2)
-    quaternion = np.array(
-        [s * rotation_axis[0], s * rotation_axis[1], s * rotation_axis[2], c])
-
-    return quaternion
-
-
-def align_with_center(particles_5d, path_center):
-    """Aligns 5d particles with the center of the path
-
-    Args:
-        particles_5d ([type]): The 5d particles
-        path_center ([type]): The center of the path
-
-    Returns:
-        [type]: The particles with their orientations aligned with the center of the path
-    """
-    center_aligned_particles = deepcopy(particles_5d)
-    number_of_cameras = int(len(particles_5d)/5)
-    for i in range(len(particles_5d[0])):
-        for j in range(number_of_cameras):
-
-            angles = particles_5d[j*5:j*5+2, i]
-            position = spherical_to_cartesian(
-                camera_radius, angles[0], angles[1])
-            direction = path_center-position
-
-            quaternion = quat_from_direction(direction)
-            center_aligned_particles[j*5+2:j*5+5,
[…]

### 2024-07-24 code/paper_setup_visualization.py

@@ -1,26 +1,0 @@
-from paper_experiment import CustomEnv
-import pybullet as p
-import pybullet_industrial as pi
-import numpy as np
-
-if __name__ == "__main__":
-
-    env = CustomEnv(True, 200)
-
-    focal_length = 0.1
-    intrinsic_matrix = [0.0]*9
-    intrinsic_matrix[0*3+0] = focal_length
-    intrinsic_matrix[1*3+1] = focal_length
-    intrinsic_matrix[2*3+2] = float(1)
-    intrinsic_matrix = tuple(intrinsic_matrix)
-    camera_state = np.array([0.0, 1, 0.6, np.pi/2, 0.0, 0.0])
-    camera_configs = []
-    pi.draw_coordinate_system(camera_state[:3],
-                              p.getQuaternionFromEuler(camera_state[3:6]))
-    camera_configs.append([camera_state[:3], camera_state[3:6]])
-
-    env.configure_cameras([camera_configs], [[
-                          intrinsic_matrix]*len(camera_configs)])
-
-    while True:
-        env.run_simulation()

### 2024-07-24 code/singular_orientation_exp.py

@@ -1,41 +1,0 @@
-import numpy as np
-from paper_experiment import CustomEnv
-
-
-if __name__ == "__main__":
-    env = CustomEnv(False)
-
-    focal_length = 0.1
-    intrinsic_matrix = [0.0]*9
-    intrinsic_matrix[0*3+0] = focal_length
-    intrinsic_matrix[1*3+1] = focal_length
-    intrinsic_matrix[2*3+2] = float(1)
-    intrinsic_matrix = tuple(intrinsic_matrix)
-
-    base_camera_state = [[0, 0, 0], [0, 0, 0]]
-
-    rotating_camera_state = [[1, 0, 0], [0, 0, 0]]
-
-    steps = 100
-
-    # create an array of angles from -pi/2 to pi/2  in n steps
-    camera_angles = np.linspace(-np.pi/2, np.pi/2, steps)
-
-    singular_values = []
-
-    for camera_angle in camera_angles:
-        rotating_camera_state[1][1] = camera_angle
-
-        singular_value = env.configure_cameras([[base_camera_state, rotating_camera_state]], [[
-            intrinsic_matrix, intrinsic_matrix]])
-
-        singular_values.append(singular_value)
-
-    env.close()
-
-    import matplotlib.pyplot as plt
-    plt.plot(camera_angles, singular_values)
-    plt.title("Singular values of the camera configurations")
-    plt.xlabel("Camera angle relative to base camera [rad]")
-    plt.ylabel("Singular values")
-    plt.show()

### 2024-07-24 code/test_camera_model.py

@@ -1,63 +1,0 @@
-import os
-
-import numpy as np
-import pybullet as p
-import pybullet_industrial as pi
-
-
-if __name__ == "__main__":
-    # setup a pybullet simulation
-    p.connect(p.GUI)
-
-    # pybullet load marker.urdf
-    dirname = os.path.dirname(__file__)
-    urdf_file = os.path.join(dirname, 'marker.urdf')
-    # load urdf
-    p.loadURDF(urdf_file, [0.0, 0, 0], [0, 0, 0, 1], useFixedBase=True)
-
-    camera_parameters = {}
-    fov = 60
-    img_width = 640
-    img_height = 480
-    near_plane = 0.1
-    far_plane = 100
-    aspect_ratio = 1
-    focal_length = 10
-
-    projection_matrix = p.computeProjectionMatrixFOV(
-        fov, aspect_ratio, near_plane, far_plane)
-
-    first_camera_pos = np.array([-0., 0, -1])
-    second_camera_pos = np.array([-0., 0., -1])
-
-    first_camera_rot = [0, 0, 0, 1]
-    second_camera_rot = p.getQuaternionFromEuler([0, 0, 0])
-
-    pi.draw_coordinate_system(first_camera_pos, first_camera_rot)
-    pi.draw_coordinate_system(second_camera_pos, second_camera_rot)
-
-    # convert quaternion to rotation matrix
-    first_camera_rot_mat = p.getMatrixFromQuaternion(first_camera_rot)
-    first_camera_rot_mat = np.array(first_camera_rot_mat).reshape(3, 3)
-    second_camera_rot_mat = p.getMatrixFromQuaternion(second_camera_rot)
-    second_camera_rot_mat = np.array(second_camera_rot_mat).reshape(3, 3)
-
-    # calculate camera target position in world frame
-    first_camera_target_pos = first_camera_pos+focal_length * \
-        first_camera_rot_mat.dot(np.array([0, 0, 1]))
-    second_camera_target_pos = second_camera_pos+focal_length * \
-        second_camera_rot_mat.dot(np.array([0, 0, 1]))
-
-    # Setup the view matrices for both camera
-    first_camera_view_matrix = p.computeViewMatrix(
-        first_camera_pos, first_camera_target_pos, np.array([0, 1, 0]))
-    second_camera_view_matrix = p.computeViewMatrix(
-        second_camera_pos, second_camera_target_pos, np.array([0, 1, 0]))
-
-    first_image = p.getCameraImage(
[…]

### 2024-07-24 code/optical_marker.py
started rewriting markers to active laser tracker setup

@@ -21,149 +21,89 @@ def compute_angle_between_vectors(v1, v2):
     return np.arccos(np.clip(np.dot(v1_u, v2_u), -1.0, 1.0))
 
 
-class OpticalMarker(pi.EndeffectorTool):
-    """Class which represents an optical marker
+class ActiveMarker(pi.EndeffectorTool):
+    """Class which represents an active marker for a laser tracker.
     Args:
         urdf_model (str): Path to the urdf model of the tool
         start_position (list): Position of the tool in the world frame
         start_orientation (list): Orientation of the tool in the world frame
         marker_positions (list): Positions of the markers in the tool frame
+        marker_orientations (list): Orientations of the markers in the tool frame
         coupled_robots (list): List of robots which are coupled to the tool
         tcp_frame (list): Position of the tool center point in the tool frame
         connector_frames (list): List of positions of the connectors in the tool frame
     """
 
     def __init__(self, urdf_model: str, start_position, start_orientation, marker_positions,
-                 coupled_robots=None, tcp_frame=None, connector_frames=None):
+                 marker_orientations, coupled_robots=None, tcp_frame=None, connector_frames=None):
         super().__init__(urdf_model, start_position, start_orientation,
                          coupled_robots, tcp_frame, connector_frames)
 
         self.marker_positions = marker_positions
-        steps = 50
-        particle_size = 0.03
-        width = particle_size*500
-        theta2 = np.linspace(-np.pi,  0, steps)
-        phi2 = np.linspace(0,  5 * 2*np.pi, steps)
+        self.marker_orientations = marker_orientations
 
-        for positions in marker_positions:
-            x_coord = particle_size * \
-                np.sin(theta2) * np.cos(phi2) + positions[0]
-            y_coord = particle_size * \
-                np.sin(theta2) * np.sin(phi2) + positions[1]
-            z_coord = particle_size * \
-                np.cos(theta2) + positions[2]
-
-            path = np.array([x_coord, y_coord, z_coord])
-            path_steps = len(path[0])
-            for i in range(1, path_steps):
-                current_point = path[:, i]
-                previous_point = path[:, i-1]
-                p.addUserDebugLine(current_point, previous_point,
-                                   lineColorRGB=[20/255, 68/255, 102/255],
-                                   lineWidth=width,
-                                   lifeTime=0,
-                                   parentObjectUniqueId=self.urdf,
[…]

### 2024-07-24 code/optical_marker.py
added a line of sight calculation back in

@@ -59,18 +59,32 @@ class ActiveMarker(pi.EndeffectorTool):
 
         current_marker_poses, current_marker_orientations = self.get_marker_poses_orientations()
 
-        for i in range(num_markers):
-            for j in range(num_cameras):
-                ray_vector = np.array(current_marker_poses[i]) - np.array(camera_positions[j])
+        ray_start_pos = []
+        ray_end_pos = []
+        for camera in camera_positions:
+            for marker_pos in current_marker_poses:
+                ray_start_pos.append(camera)
+                ray_end_pos.append(marker_pos)
 
-                marker_rotation = p.getMatrixFromQuaternion(current_marker_orientations[i])
+        ray_intersections = p.rayTestBatch(ray_start_pos, ray_end_pos)
+
+        for i in range(num_cameras):
+            for j in range(num_markers):
+                ray_index = i * num_markers + j
+                # Check if there is a clear line of sight
+                if ray_intersections[ray_index][0] != -1:
+                    continue
+
+                ray_vector = np.array(current_marker_poses[j]) - np.array(camera_positions[i])
+
+                marker_rotation = p.getMatrixFromQuaternion(current_marker_orientations[j])
                 marker_rotation = np.array(marker_rotation).reshape(3, 3)
                 marker_vector = marker_rotation[:, 2]
 
                 angle = compute_angle_between_vectors(ray_vector, marker_vector)
 
-                if angle < field_of_views[i] / 2:
-                    visibility[j][i] = 1
+                if angle < field_of_views[j] / 2:
+                    visibility[i][j] = 1
 
         return visibility
 

### 2024-07-24 code/optical_marker.py
fixed ray direction to come from marker for proper angle computation

@@ -75,7 +75,7 @@ class ActiveMarker(pi.EndeffectorTool):
                 if ray_intersections[ray_index][0] != -1:
                     continue
 
-                ray_vector = np.array(current_marker_poses[j]) - np.array(camera_positions[i])
+                ray_vector = np.array(camera_positions[i]) - np.array(current_marker_poses[j])
 
                 marker_rotation = p.getMatrixFromQuaternion(current_marker_orientations[j])
                 marker_rotation = np.array(marker_rotation).reshape(3, 3)

### 2024-07-24 code/optical_marker.py
fixed tests to make sure new marker node is running

@@ -42,27 +42,27 @@ class ActiveMarker(pi.EndeffectorTool):
         self.marker_positions = marker_positions
         self.marker_orientations = marker_orientations
 
-    def compute_visibility(self, camera_positions, field_of_views):
+    def compute_visibility(self, tracker_positions, field_of_views):
         """ Computes the visibility of the markers given a laser position and marker orientations.
 
         Args:
-            camera_positions (list): List of laser tracker positions
+            tracker_positions (list): List of laser tracker positions
             field_of_views (list): List of field of views of the markers
 
         Returns:
             array: A matrix of size (number of markers, number of cameras)
                    which contains the visibility of the markers for each camera
         """
-        num_cameras = len(camera_positions)
+        num_cameras = len(tracker_positions)
         num_markers = len(self.marker_positions)
         visibility = np.zeros((num_cameras, num_markers))
 
-        current_marker_poses, current_marker_orientations = self.get_marker_poses_orientations()
+        current_marker_positions, current_marker_orientations = self.get_marker_poses_orientations()
 
         ray_start_pos = []
         ray_end_pos = []
-        for camera in camera_positions:
-            for marker_pos in current_marker_poses:
+        for camera in tracker_positions:
+            for marker_pos in current_marker_positions:
                 ray_start_pos.append(camera)
                 ray_end_pos.append(marker_pos)
 
@@ -75,13 +75,15 @@ class ActiveMarker(pi.EndeffectorTool):
                 if ray_intersections[ray_index][0] != -1:
                     continue
 
-                ray_vector = np.array(camera_positions[i]) - np.array(current_marker_poses[j])
+                ray_vector = np.array(tracker_positions[i]) - np.array(current_marker_positions[j])
 
                 marker_rotation = p.getMatrixFromQuaternion(current_marker_orientations[j])
                 marker_rotation = np.array(marker_rotation).reshape(3, 3)
                 marker_vector = marker_rotation[:, 2]
 
+
                 angle = compute_angle_between_vectors(ray_vector, marker_vector)
+
 
                 if angle < field_of_views[j] / 2:
                     visibility[i][j] = 1
@@ -135,8 +137,8 @@ class TestActiveMarker(unittest.TestCase):
                                   [0, 0, 0, 1],
                                   [0, 0, 0, 1]])
 
[…]

### 2024-07-24 code/optical_marker.py
adjusted paths and added crude run_simulation routing

@@ -115,6 +115,16 @@ class ActiveMarker(pi.EndeffectorTool):
 
         return marker_positions, marker_orientations
 
+    def set_optical_system_parameters(self,particles):
+        """Convenient interface to set marker positions and orientations all in one go.
+
+        Args:
+            particles (list): List of markers where each marker is a
+                              list containing the position and orientation in a 6d vector
+        """
+        self.marker_positions = [particle[:3] for particle in particles]
+        self.marker_orientations = [particle[3:] for particle in particles]
+
 
 class TestActiveMarker(unittest.TestCase):
 

### 2024-07-24 code/paper_experiment.py

@@ -5,7 +5,7 @@ import pybullet as p
 import pybullet_industrial as pi
 
 
-from optical_marker import OpticalMarker
+from optical_marker import ActiveMarker
 from particle_optimizer import particle_optimizer, particle_optimizer_log
 
 
@@ -15,7 +15,7 @@ from particle_optimizer import particle_optimizer, particle_optimizer_log
 
 class LaserTrackerEnv:
 
-    def __init__(self,  rendering=False, steps=20):
+    def __init__(self,  rendering=False, steps=20,path=None):
         if rendering is False:
             self.physics_client = p.connect(p.DIRECT)
         else:
@@ -23,11 +23,21 @@ class LaserTrackerEnv:
                                             '--background_color_green=1 ' +
                                             '--background_color_blue=1')
             p.configureDebugVisualizer(p.COV_ENABLE_GUI, 0)
+
         p.configureDebugVisualizer(p.COV_ENABLE_RENDERING, 0)
         p.setGravity(0, 0, -9.81)
         p.setPhysicsEngineParameter(numSolverIterations=1000)
         self._time_step_length = 0.1
         p.setTimeStep(self._time_step_length)
+
+        if path is None:
+            target_position = np.array([1.9, 0, 1.03])
+            steps = 100
+            self.path = pi.build_box_path(
+                target_position, [0.5, 0.6], 0.1, [0, 0, 0, 1], steps)
+
+        else:
+            self.path = path
 
 
         dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects')
@@ -68,20 +78,22 @@ class LaserTrackerEnv:
         self.second_robot = pi.RobotBase(comau_urdf, start_pos2, start_orientation2)
 
 
-        self.marker = OpticalMarker(os.path.join(dirname, 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1],
-                                 [[0.1, 0.1, 0.1],
+        self.marker = ActiveMarker(os.path.join(os.path.dirname(__file__), 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1],
+                                  [[0.1, 0.1, 0.1],
                                   [0.1, -0.1, 0.1],
                                   [-0.1, 0.1, 0.1],
-                                  [-0.1, -0.1, 0.1]])
+                                  [-0.1, -0.1, 0.1]],
+                                 [[0, 0, 0, 1],
+                                  [0, 0, 0, 1],
[…]

### 2024-07-24 code/paper_experiment.py
converted euler angles to quaternion

@@ -31,13 +31,16 @@ class LaserTrackerEnv:
         p.setTimeStep(self._time_step_length)
 
         if path is None:
-            target_position = np.array([1.9, 0, 1.03])
+            target_position = np.array([1.9, 0, 1.03])+np.array([-3.6353162, -0.6, 0])
             steps = 100
             self.path = pi.build_box_path(
                 target_position, [0.5, 0.6], 0.1, [0, 0, 0, 1], steps)
 
         else:
             self.path = path
+
+
+        self.path.draw()
 
 
         dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects')
@@ -113,7 +116,8 @@ class LaserTrackerEnv:
 
         # extract the marker poses and laser tracker positions from the particles
         laser_tracker_positions = [particle[:3] for particle in particles]
-        marker_poses = [particle[3:] for particle in particles]
+        #extract the marker poses and convert the euler angles to quaternions
+        marker_poses = [np.append(particle[3:6],p.getQuaternionFromEuler(particle[6:9])) for particle in particles]
 
         field_of_views = [np.pi/4]* len(marker_poses)
 
@@ -123,7 +127,7 @@ class LaserTrackerEnv:
         # adding it to the visibility index
 
         for _ in range(20):
-            self.marker.set_tool_pose(*self.marker.get_start_pose())
+            self.marker.set_tool_pose(*self.path.get_start_pose())
             for _ in range(50):
                 p.stepSimulation()
 
@@ -216,4 +220,5 @@ if __name__ == "__main__":
     env = LaserTrackerEnv(rendering=True)
 
     while True:
-        p.stepSimulation()
+        env.run_simulation([np.random.rand(9)*0.3*np.pi]*10)
+

### 2024-07-24 code/optical_marker.py
added visualization for lines of sight

@@ -67,11 +67,19 @@ class ActiveMarker(pi.EndeffectorTool):
                 ray_end_pos.append(marker_pos)
 
         ray_intersections = p.rayTestBatch(ray_start_pos, ray_end_pos)
+        # draw each ray using p.addUserDebugLine
+        for i in range(len(ray_start_pos)):
+            for j in range(len(ray_end_pos)):
+                if ray_intersections[i][0] != -1:
+                    p.addUserDebugLine(ray_start_pos[i], ray_end_pos[j], [0, 0, 1], 1)
+                else:
+                    p.addUserDebugLine(ray_start_pos[i], ray_end_pos[j], [1, 0, 0], 1)
+
 
         for i in range(num_cameras):
             for j in range(num_markers):
                 ray_index = i * num_markers + j
-                # Check if there is a clear line of sight
+
                 if ray_intersections[ray_index][0] != -1:
                     continue
 

### 2024-07-24 code/paper_experiment.py

@@ -31,8 +31,8 @@ class LaserTrackerEnv:
         p.setTimeStep(self._time_step_length)
 
         if path is None:
-            target_position = np.array([1.9, 0, 1.03])+np.array([-3.6353162, -0.6, 0])
-            steps = 100
+            target_position = np.array([1.6, 0, 1.03])+np.array([-3.6353162, -0.6, 0])
+            steps = 30
             self.path = pi.build_box_path(
                 target_position, [0.5, 0.6], 0.1, [0, 0, 0, 1], steps)
 
@@ -90,11 +90,11 @@ class LaserTrackerEnv:
                                   [0, 0, 0, 1],
                                   [0, 0, 0, 1],
                                   [0, 0, 0, 1]])
-        self.marker.couple(self.robot)
+        self.marker.couple(self.robot,'link6')
 
         self.second_endeffector = pi.EndeffectorTool(os.path.join(os.path.dirname(__file__), 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1])
 
-        self.second_endeffector.couple(self.second_robot)
+        self.second_endeffector.couple(self.second_robot,'link6')
 
         p.configureDebugVisualizer(p.COV_ENABLE_RENDERING, 1)
 
@@ -125,16 +125,22 @@ class LaserTrackerEnv:
 
         # iterate over a path and compute the visibility matrix at each step,
         # adding it to the visibility index
-
         for _ in range(20):
-            self.marker.set_tool_pose(*self.path.get_start_pose())
-            for _ in range(50):
-                p.stepSimulation()
+            joint_angles ={
+                'q1': -0.1910351778338952,
+                'q2': 0.6219868844708127,
+                'q3': -2.058245174714272,
+                'q4': -0.21271062205029018,
+                'q5': -2.1816615644867423,
+                'q6': -0.09411935083551597
+            }
+            self.robot.set_joint_position(joint_angles)
+            p.stepSimulation()
 
 
         for position,orientation,_ in self.path:
             self.marker.set_tool_pose(position,orientation)
-            for _ in range(10):
+            for _ in range(100):
                 p.stepSimulation()
             visibility_index.append(self.marker.compute_visibility(laser_tracker_positions, field_of_views))
 
@@ -220,5 +226,5 @@ if __name__ == "__main__":
     env = LaserTrackerEnv(rendering=True)
 
     while True:
[…]

### 2024-07-24 code/paper_experiment.py
added offset to test position to see wheter the visualisation works

@@ -226,5 +226,5 @@ if __name__ == "__main__":
     env = LaserTrackerEnv(rendering=True)
 
     while True:
-        env.run_simulation([[-4,-1,3,0,0,0,0,0,0],[-3,-3,1.2,0,0,0,0,0,0]])
+        env.run_simulation([[-4,-1,3,0,0,0.3,0,0,0],[-3,-3,1.2,0,-0.3,0.1,0,0,0]])
 

### 2024-07-24 code/optical_marker.py
adjusted optical marker so that only 1-1 correspondance of marker and tracker is used for visibility

@@ -21,6 +21,7 @@ def compute_angle_between_vectors(v1, v2):
     return np.arccos(np.clip(np.dot(v1_u, v2_u), -1.0, 1.0))
 
 
+
 class ActiveMarker(pi.EndeffectorTool):
     """Class which represents an active marker for a laser tracker.
     Args:
@@ -29,72 +30,57 @@ class ActiveMarker(pi.EndeffectorTool):
         start_orientation (list): Orientation of the tool in the world frame
         marker_positions (list): Positions of the markers in the tool frame
         marker_orientations (list): Orientations of the markers in the tool frame
+        tracker_positions (list): Positions of the laser trackers
         coupled_robots (list): List of robots which are coupled to the tool
         tcp_frame (list): Position of the tool center point in the tool frame
         connector_frames (list): List of positions of the connectors in the tool frame
     """
 
     def __init__(self, urdf_model: str, start_position, start_orientation, marker_positions,
-                 marker_orientations, coupled_robots=None, tcp_frame=None, connector_frames=None):
+                 marker_orientations, tracker_positions, coupled_robots=None, tcp_frame=None, connector_frames=None):
         super().__init__(urdf_model, start_position, start_orientation,
                          coupled_robots, tcp_frame, connector_frames)
 
         self.marker_positions = marker_positions
         self.marker_orientations = marker_orientations
+        self.tracker_positions = tracker_positions
 
-    def compute_visibility(self, tracker_positions, field_of_views):
+    def compute_visibility(self, field_of_views):
         """ Computes the visibility of the markers given a laser position and marker orientations.
 
         Args:
-            tracker_positions (list): List of laser tracker positions
             field_of_views (list): List of field of views of the markers
 
         Returns:
-            array: A matrix of size (number of markers, number of cameras)
-                   which contains the visibility of the markers for each camera
+            list: A list which contains the visibility of each marker for its corresponding tracker
         """
-        num_cameras = len(tracker_positions)
         num_markers = len(self.marker_positions)
-        visibility = np.zeros((num_cameras, num_markers))
+        visibility = np.zeros(num_markers)
 
         current_marker_positions, current_marker_orientations = self.get_marker_poses_orientations()
 
-        ray_start_pos = []
[…]

### 2024-07-24 code/optical_marker.py
adjusted experiment to new interface

@@ -112,13 +112,10 @@ class ActiveMarker(pi.EndeffectorTool):
     def set_optical_system_parameters(self,particles):
         """Convenient interface to set marker positions and orientations all in one go.
 
-        Args:
-            particles (list): List of markers where each marker is a
-                              list containing the position and orientation in a 6d vector
         """
         self.tracker_positions = [particle[:3] for particle in particles]
         self.marker_positions = [particle[3:6] for particle in particles]
-        self.marker_orientations = [particle[6:] for particle in particles]
+        self.marker_orientations = [p.getQuaternionFromEuler(particle[6:9]) for particle in particles]
 
 class TestActiveMarker(unittest.TestCase):
 

### 2024-07-24 code/paper_experiment.py

@@ -89,7 +89,10 @@ class LaserTrackerEnv:
                                  [[0, 0, 0, 1],
                                   [0, 0, 0, 1],
                                   [0, 0, 0, 1],
-                                  [0, 0, 0, 1]])
+                                  [0, 0, 0, 1]],
+                                  [[0.0, 0.0, 0.5],
+                                   [0.0, 0.0, 0.6]])
+
         self.marker.couple(self.robot,'link6')
 
         self.second_endeffector = pi.EndeffectorTool(os.path.join(os.path.dirname(__file__), 'marker.urdf'), [0, 0, 0], [0, 0, 0, 1])
@@ -121,7 +124,7 @@ class LaserTrackerEnv:
 
         field_of_views = [np.pi/4]* len(marker_poses)
 
-        self.marker.set_optical_system_parameters(marker_poses)
+        self.marker.set_optical_system_parameters(particles)
 
         # iterate over a path and compute the visibility matrix at each step,
         # adding it to the visibility index
@@ -142,7 +145,7 @@ class LaserTrackerEnv:
             self.marker.set_tool_pose(position,orientation)
             for _ in range(100):
                 p.stepSimulation()
-            visibility_index.append(self.marker.compute_visibility(laser_tracker_positions, field_of_views))
+            visibility_index.append(self.marker.compute_visibility( field_of_views))
 
         return visibility_index
 

### 2024-07-24 code/optical_marker.py
folded field of view into marker class

@@ -37,24 +37,24 @@ class ActiveMarker(pi.EndeffectorTool):
     """
 
     def __init__(self, urdf_model: str, start_position, start_orientation, marker_positions,
-                 marker_orientations, tracker_positions, coupled_robots=None, tcp_frame=None, connector_frames=None):
+                 marker_orientations, tracker_positions,
+                 marker_fov,coupled_robots=None, tcp_frame=None, connector_frames=None):
         super().__init__(urdf_model, start_position, start_orientation,
                          coupled_robots, tcp_frame, connector_frames)
 
         self.marker_positions = marker_positions
         self.marker_orientations = marker_orientations
         self.tracker_positions = tracker_positions
+        self.marker_fov = marker_fov
 
-    def compute_visibility(self, field_of_views):
+    def compute_visibility(self):
         """ Computes the visibility of the markers given a laser position and marker orientations.
-
-        Args:
-            field_of_views (list): List of field of views of the markers
 
         Returns:
             list: A list which contains the visibility of each marker for its corresponding tracker
         """
         num_markers = len(self.marker_positions)
+        field_of_views = [self.marker_fov]*num_markers
         visibility = np.zeros(num_markers)
 
         current_marker_positions, current_marker_orientations = self.get_marker_poses_orientations()
@@ -131,9 +131,9 @@ class TestActiveMarker(unittest.TestCase):
                                  [[0, 0, 0, 1],
                                   [0, 0, 0, 1]],
                                   [[0.0, 0.0, 0.5],
-                                   [0.0, 0.0, 0.6]])
+                                   [0.0, 0.0, 0.6]],
+                                   np.pi/4)
 
-        self.field_of_views = [np.pi/4 , np.pi/4]
 
     def tearDown(self):
         """Clean up after the test."""
@@ -143,7 +143,7 @@ class TestActiveMarker(unittest.TestCase):
         """Test visibility for a marker directly in front of the laser tracker."""
         self.tool.marker_orientations = [p.getQuaternionFromEuler([0, 0,0]),
                                             p.getQuaternionFromEuler([0, 0,0])]
-        visibility = self.tool.compute_visibility( self.field_of_views)
+        visibility = self.tool.compute_visibility( )
         # The first camera should see all markers
         print("visibility: ",visibility)
         self.assertEqual(np.sum(visibility), 2)
[…]

### 2024-07-24 code/paper_experiment.py

@@ -91,7 +91,8 @@ class LaserTrackerEnv:
                                   [0, 0, 0, 1],
                                   [0, 0, 0, 1]],
                                   [[0.0, 0.0, 0.5],
-                                   [0.0, 0.0, 0.6]])
+                                   [0.0, 0.0, 0.6]],
+                                   np.pi/4)
 
         self.marker.couple(self.robot,'link6')
 
@@ -117,13 +118,6 @@ class LaserTrackerEnv:
         visibility_index = []
         self.reset()
 
-        # extract the marker poses and laser tracker positions from the particles
-        laser_tracker_positions = [particle[:3] for particle in particles]
-        #extract the marker poses and convert the euler angles to quaternions
-        marker_poses = [np.append(particle[3:6],p.getQuaternionFromEuler(particle[6:9])) for particle in particles]
-
-        field_of_views = [np.pi/4]* len(marker_poses)
-
         self.marker.set_optical_system_parameters(particles)
 
         # iterate over a path and compute the visibility matrix at each step,
@@ -145,7 +139,7 @@ class LaserTrackerEnv:
             self.marker.set_tool_pose(position,orientation)
             for _ in range(100):
                 p.stepSimulation()
-            visibility_index.append(self.marker.compute_visibility( field_of_views))
+            visibility_index.append(self.marker.compute_visibility( ))
 
         return visibility_index
 
@@ -228,6 +222,7 @@ def run_experiment():
 if __name__ == "__main__":
     env = LaserTrackerEnv(rendering=True)
 
+    particles = [[-4,-1,3,0,0,0.3,0,0,0],[-3,-3,1.2,0,-0.3,0.1,0,0,0]]
     while True:
-        env.run_simulation([[-4,-1,3,0,0,0.3,0,0,0],[-3,-3,1.2,0,-0.3,0.1,0,0,0]])
+        env.run_simulation(particles)
 

### 2024-07-24 code/paper_experiment.py
added constraints and costs for first working version

@@ -5,6 +5,8 @@ import pybullet as p
 import pybullet_industrial as pi
 
 
+from rich.console import Console
+
 from optical_marker import ActiveMarker
 from particle_optimizer import particle_optimizer, particle_optimizer_log
 
@@ -118,7 +120,7 @@ class LaserTrackerEnv:
         visibility_index = []
         self.reset()
 
-        self.marker.set_optical_system_parameters(particles)
+        self.marker.set_optical_system_parameters(particles.T)
 
         # iterate over a path and compute the visibility matrix at each step,
         # adding it to the visibility index
@@ -141,6 +143,9 @@ class LaserTrackerEnv:
                 p.stepSimulation()
             visibility_index.append(self.marker.compute_visibility( ))
 
+        visibility_index = np.array(visibility_index)
+
+        visibility_index = visibility_index.T
         return visibility_index
 
     def set_timestep_length(self, time_step_length):
@@ -164,39 +169,41 @@ class LaserTrackerEnv:
 
 
 def run_experiment():
-
+    console = Console()
     def objective_function(particles):
 
-        # divide the particles into list of camera states
-        camera_particles = np.split(particles, len(particles)/6, axis=0)
+        visibility_cost = env.run_simulation(particles)
+        visibility_cost_sum = np.sum(visibility_cost,axis=1)
 
-        camera_states = []
-        for i in range(len(camera_particles[0][0])):
-            camera_state = [[state[:3, i], state[3:, i]]
-                            for state in camera_particles]
-            camera_states.append(camera_state)
 
-        focal_length = 0.1
-        intrinsic_matrix = [0.0]*9
-        intrinsic_matrix[0*3+0] = focal_length
-        intrinsic_matrix[1*3+1] = focal_length
-        intrinsic_matrix[2*3+2] = float(1)
-        intrinsic_matrix = tuple(intrinsic_matrix)
+        #constraint costs
 
-        env.configure_cameras(
-            camera_states, [[intrinsic_matrix]*len(camera_particles)]*len(particles[0]))
-        costs = env.run_simulation()
-        costs = np.sum(costs, axis=0)
+        #constraints keeping the height of the laser tracker betwen 0.8 and 2
+        height_constraint = np.clip(np.abs(particles[2,:]-1.03)-0.8,0,-1e6)
 
-        # penalize the cameras that are below the ground
[…]

### 2024-07-24 code/paper_experiment.py
fixed constraint calculation

@@ -179,7 +179,8 @@ def run_experiment():
         #constraint costs
 
         #constraints keeping the height of the laser tracker betwen 0.8 and 2
-        height_constraint = np.clip(np.abs(particles[2,:]-1.03)-0.8,0,-1e6)
+        height_constraint = np.array([ -1e6 if x < 0.8 or x > 2 else 0 for x in particles[2,:]])
+
 
 
 
@@ -188,8 +189,8 @@ def run_experiment():
         cylinder_height = 0.4
         marker_positions = particles[3:6,:]
 
-        radius_constraint = np.clip(np.linalg.norm(marker_positions[:2,:],axis=0)-cylinder_radius,0,-1e6)
-        height_constraint = np.clip(np.abs(marker_positions[2,:])-cylinder_height,0,-1e6)
+        radius_constraint = np.array([ -1e6 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
+        height_constraint = np.array([ -1e6 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
 
 
 
@@ -206,19 +207,30 @@ def run_experiment():
     env = LaserTrackerEnv()
 
     n_particles = 100
-    initial_population = np.random.rand(
-        9, n_particles)*0.3*np.pi
+
+
+    initial_tracker_positions = np.random.rand(3, n_particles)*2
+    initial_marker_positions = np.random.rand(3, n_particles)*0.6
+    initial_marker_orientations = np.random.rand(3, n_particles)*2*np.pi
+
+    initial_population = np.vstack(
+        [initial_tracker_positions, initial_marker_positions, initial_marker_orientations])
+
 
 
     np.save("initial_states.npy",initial_population)
 
-    start_velocities = np.random.randn(
-        9, n_particles) * 0.5
+    start_tracker_velocities = np.random.rand(3, n_particles)*0.4
+    start_marker_velocities = np.random.rand(3, n_particles)*0.01
+    start_orientations_velocities = np.random.rand(3, n_particles)*0.1
+
+    start_velocities = np.vstack(
+        [start_tracker_velocities, start_marker_velocities, start_orientations_velocities])
 
     with console.status("Finding optimal camera placements"):
         particle_log, obj_log = particle_optimizer_log(objective_function, initial_population,
                                                         start_velocities, max_iter=100)
-    env.close()
+
[…]

### 2024-07-24 code/paper_experiment.py
switched sign and produced first usable results

@@ -173,13 +173,13 @@ def run_experiment():
     def objective_function(particles):
 
         visibility_cost = env.run_simulation(particles)
-        visibility_cost_sum = np.sum(visibility_cost,axis=1)
+        visibility_cost_sum = -1*np.sum(visibility_cost,axis=1)
 
 
         #constraint costs
 
         #constraints keeping the height of the laser tracker betwen 0.8 and 2
-        height_constraint = np.array([ -1e6 if x < 0.8 or x > 2 else 0 for x in particles[2,:]])
+        height_constraint = np.array([ 1e6 if x < 0.8 or x > 2 else 0 for x in particles[2,:]])
 
 
 
@@ -189,8 +189,8 @@ def run_experiment():
         cylinder_height = 0.4
         marker_positions = particles[3:6,:]
 
-        radius_constraint = np.array([ -1e6 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
-        height_constraint = np.array([ -1e6 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
+        radius_constraint = np.array([ 1e6 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
+        height_constraint = np.array([ 1e6 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
 
 
 

### 2024-07-24 code/paper_visualization.py
added proper visualisation

@@ -1,42 +1,22 @@
-from paper_experiment import CustomEnv
+from paper_experiment import LaserTrackerEnv
 import pybullet as p
 import pybullet_industrial as pi
 import numpy as np
 
 if __name__ == "__main__":
-    camera_radius = 2.5
-    optimal_states = np.load("optimal_states_6_cameras.npy")
-    optimal_camera_states = np.split(
-        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
+    particle_log = np.load("particle_log.npy")
+    objective_log = np.load("objective_log.npy")
+    env = LaserTrackerEnv(rendering=True)
 
-    initial_states = np.load("initial_states_6_cameras.npy")
-    initial_camera_states = np.split(
-        initial_states, len(initial_states)/6, axis=0)
-    env = CustomEnv(True)
+    last_particle_round = particle_log[-1]
 
-    focal_length = 0.1
-    intrinsic_matrix = [0.0]*9
-    intrinsic_matrix[0*3+0] = focal_length
-    intrinsic_matrix[1*3+1] = focal_length
-    intrinsic_matrix[2*3+2] = float(1)
-    intrinsic_matrix = tuple(intrinsic_matrix)
+    best_particle = last_particle_round[:,np.argmin(objective_log[-1])]
 
-    camera_configs = []
-    for camera_state in optimal_camera_states:
-        pi.draw_coordinate_system(camera_state[:3],
-                                  p.getQuaternionFromEuler(camera_state[3:6]))
-        camera_configs.append([camera_state[:3], camera_state[3:6]])
+    print("Best particle: ", best_particle)
 
-    env.configure_cameras([camera_configs], [[
-                          intrinsic_matrix]*len(camera_configs)])
+    # reshape to 9x1
+    best_particle = best_particle.reshape(9,1)
 
-    for camera_state in initial_camera_states:
-        for i in range(len(camera_state[0])):
-            pi.draw_coordinate_system(camera_state[:3, i],
-                                      p.getQuaternionFromEuler(
-                                          camera_state[3:6, i]),
-                                      length=0.05, width=1.0)
-    env._compute_visibility()
-    env.run_simulation()
     while True:
-        p.stepSimulation()
+        env.run_simulation(best_particle)
+

### 2024-07-24 code/multi_hist_comparison.py
added evaluation and decreased penalty term for prettier plots

@@ -1,141 +1,0 @@
-import matplotlib
-import numpy as np
-import pandas as pd
-import seaborn as sns
-import matplotlib.pyplot as plt
-
-from paper_experiment import CustomEnv
-from paper_evaluation import DirectVisEnv, OrientationVisEnv, get_fov_from_intrinsic_matrix
-
-kit_deep_blue = "#144466"
-kit_red = "#B2372C"
-
-kit_green = "#009682"
-kit_yellow = "#EEB70D"
-
-light_blue = "#8EB1C7"
-pink = "#FFC6D9"
-purble = "#A288A6"
-shining_blue = "#70D6FF"
-
-colors = [kit_deep_blue, kit_green, kit_yellow,
-          light_blue, pink, purble, shining_blue]
-
-
-def load_camera_data(number_of_cameras):
-    optimal_states = np.load(
-        "optimal_states_"+str(number_of_cameras)+"_cameras.npy")
-    optimal_camera_states = np.split(
-        optimal_states.flatten(), len(optimal_states.flatten())/6, axis=0)
-
-    initial_states = np.load(
-        "initial_states_"+str(number_of_cameras)+"_cameras.npy")
-    initial_camera_states = np.split(
-        initial_states, len(initial_states)/6, axis=0)
-
-    focal_length = 0.1
-    intrinsic_matrix = [0.0]*9
-    intrinsic_matrix[0*3+0] = focal_length
-    intrinsic_matrix[1*3+1] = focal_length
-    intrinsic_matrix[2*3+2] = float(1)
-    intrinsic_matrix = tuple(intrinsic_matrix)
-
-    optimal_camera_configs = []
-    for camera_state in optimal_camera_states:
-        optimal_camera_configs.append([camera_state[:3], camera_state[3:6]])
-
-    initial_camera_configs = []
-    for i in range(len(initial_camera_states[0][0])):
-        initial_camera_configs.append([
-            [init_state[:3, i], init_state[3:6, i]] for init_state in initial_camera_states])
-
-    return optimal_camera_configs, initial_camera_configs, intrinsic_matrix
-
-
-def get_triangulation_data(number_of_cameras):
-    optimal_camera_configs, initial_camera_configs, intrinsic_matrix = load_camera_data(
-        number_of_cameras)
-
-    env = CustomEnv(False)
-    optimal_triangulation_error = env.configure_cameras([optimal_camera_configs], [[
-        intrinsic_matrix]*len(optimal_camera_configs)])
[…]

### 2024-07-24 code/paper_evaluation.py

@@ -1,0 +1,33 @@
+import numpy as np
+import seaborn as sns
+import matplotlib.pyplot as plt
+
+# Load the objective log data
+objective_log = np.load("objective_log.npy")
+
+# Calculate the mean and the percentiles for the hull curve
+mean_values = np.mean(objective_log, axis=1)
+lower_boundary = np.min(objective_log, axis=1)
+upper_boundary = np.max(objective_log, axis=1)
+
+# Set the color
+kit_deep_blue = "#144466"
+see_through_blue = "#14446680"  # Adding 80 to make it semi-transparent
+
+# Set seaborn theme to white
+sns.set_theme(style="white")
+
+# Plot the mean and the hull curve
+plt.figure(figsize=(3, 2))
+
+# Plot the mean
+sns.lineplot(x=range(objective_log.shape[0]), y=mean_values, color=kit_deep_blue, label='Mean')
+
+# Plot the hull curve
+plt.fill_between(range(objective_log.shape[0]), lower_boundary, upper_boundary, color=see_through_blue, label='5th-95th percentile range')
+
+plt.xlabel("Solver iteration")
+plt.ylabel("Objective function")
+plt.legend()
+
+plt.show()

### 2024-07-24 code/paper_experiment.py

@@ -175,11 +175,12 @@ def run_experiment():
         visibility_cost = env.run_simulation(particles)
         visibility_cost_sum = -1*np.sum(visibility_cost,axis=1)
 
+        #TODO add cost that keeps the tracker as close as possible to the surface of the robot
 
         #constraint costs
 
         #constraints keeping the height of the laser tracker betwen 0.8 and 2
-        height_constraint = np.array([ 1e6 if x < 0.8 or x > 2 else 0 for x in particles[2,:]])
+        height_constraint = np.array([ 1e2 if x < 0.8 or x > 2 else 0 for x in particles[2,:]])
 
 
 
@@ -189,8 +190,8 @@ def run_experiment():
         cylinder_height = 0.4
         marker_positions = particles[3:6,:]
 
-        radius_constraint = np.array([ 1e6 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
-        height_constraint = np.array([ 1e6 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
+        radius_constraint = np.array([ 1e2 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
+        height_constraint = np.array([ 1e2 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
 
 
 
@@ -229,7 +230,7 @@ def run_experiment():
 
     with console.status("Finding optimal camera placements"):
         particle_log, obj_log = particle_optimizer_log(objective_function, initial_population,
-                                                        start_velocities, max_iter=100)
+                                                        start_velocities, max_iter=30)
 
     # save particle log and objective log
     np.save("particle_log.npy", particle_log)

### 2024-07-24 paper/paper.tex
added deviation cost and laser tracker

@@
 The marker of course also cant be within the mesh of the manufacturing system, but since this would result in a line of sight visibility of 0 we do not need to encode this constraint explicitly.
+If we still want the marker close to the surface of the system such as when mounting them magnetically we can add a penalty function to the objective function that penalizes markers that are too far away from the flange.
+This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.
 

### 2024-07-24 code/paper_experiment.py

@@ -176,6 +176,8 @@ def run_experiment():
         visibility_cost_sum = -1*np.sum(visibility_cost,axis=1)
 
         #TODO add cost that keeps the tracker as close as possible to the surface of the robot
+        marker_positions = particles[3:6,:]
+        deviation_cost = np.linalg.norm(marker_positions, axis=0)
 
         #constraint costs
 
@@ -188,7 +190,6 @@ def run_experiment():
         #constraint keeping the marker position within a cylinder of given radius and height
         cylinder_radius = 0.5
         cylinder_height = 0.4
-        marker_positions = particles[3:6,:]
 
         radius_constraint = np.array([ 1e2 if np.linalg.norm(marker_positions[:2,i]) > cylinder_radius else 0 for i in range(marker_positions.shape[1])])
         height_constraint = np.array([ 1e2 if marker_positions[2,i] < 0 or marker_positions[2,i] > cylinder_height else 0 for i in range(marker_positions.shape[1])])
@@ -198,7 +199,7 @@ def run_experiment():
 
 
 
-        costs = visibility_cost_sum+ height_constraint + radius_constraint + height_constraint
+        costs = visibility_cost_sum+ deviation_cost+height_constraint + radius_constraint + height_constraint
 
         print(costs)
 

### 2024-07-24 code/paper_visualization.py

@@ -2,6 +2,7 @@ from paper_experiment import LaserTrackerEnv
 import pybullet as p
 import pybullet_industrial as pi
 import numpy as np
+import os
 
 if __name__ == "__main__":
     particle_log = np.load("particle_log.npy")
@@ -16,6 +17,22 @@ if __name__ == "__main__":
 
     # reshape to 9x1
     best_particle = best_particle.reshape(9,1)
+
+    laser_tracker_position = best_particle[:3].flatten()
+
+    laser_tracker_path = dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects','laser_tracker.urdf')
+    #load laser tracker at optimal position
+    #p.loadURDF(laser_tracker_path, laser_tracker_position, p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True, globalScaling=0.001)
+
+    marker_position = best_particle[3:6].flatten()
+    marker_orientation = p.getQuaternionFromEuler(best_particle[6:9])
+
+    # visualize the markers as coordinate system relative to the flange.
+    print(marker_position)
+    print(env.robot.urdf)
+    print(env.robot._convert_endeffector('link6'))
+    pi.draw_coordinate_system(marker_position,marker_orientation,
+                              parent_id=env.robot.urdf,parent_index=env.robot._convert_endeffector('link6'))
 
     while True:
         env.run_simulation(best_particle)

### 2024-07-25 code/paper_visualization.py
readded tracker to visualization

@@ -22,7 +22,7 @@ if __name__ == "__main__":
 
     laser_tracker_path = dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects','laser_tracker.urdf')
     #load laser tracker at optimal position
-    #p.loadURDF(laser_tracker_path, laser_tracker_position, p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True, globalScaling=0.001)
+    p.loadURDF(laser_tracker_path, laser_tracker_position, p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True, globalScaling=0.001)
 
     marker_position = best_particle[3:6].flatten()
     marker_orientation = p.getQuaternionFromEuler(best_particle[6:9])

### 2024-07-25 code/paper_experiment.py
adjusted paper experiment for first scenario

@@ -33,16 +33,22 @@ class LaserTrackerEnv:
         p.setTimeStep(self._time_step_length)
 
         if path is None:
-            target_position = np.array([1.6, 0, 1.03])+np.array([-3.6353162, -0.6, 0])
+            target_position = np.array([-2.4,-3,0.95])+np.array([-0.6,0.45,0.4])
             steps = 30
             self.path = pi.build_box_path(
                 target_position, [0.5, 0.6], 0.1, [0, 0, 0, 1], steps)
+            new_orientation = p.getQuaternionFromEuler([np.pi,0,0])
+            #created a array with dimension steps x 4
+            orientation_array = np.array([new_orientation for _ in range(steps)])
+            self.path.orientations=orientation_array.T
 
         else:
             self.path = path
 
 
         self.path.draw()
+
+        self.second_path = pi.linear_interpolation([-1.2, -0.6, 0.95], [-2.6, -3.4, 1.2], steps)
 
 
         dirname = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects')
@@ -65,11 +71,11 @@ class LaserTrackerEnv:
                    useFixedBase=True, globalScaling=0.001)
 
         mm_path = os.path.join(dirname,  'modules', 'milling_module.urdf')
-        p.loadURDF(mm_path, [-0.6,-4.6,0.95], p.getQuaternionFromEuler([0,0,np.pi/2]),
+        p.loadURDF(mm_path, [-2.4,-3,0.95], p.getQuaternionFromEuler([0,0,np.pi/2]),
                    useFixedBase=True, globalScaling=0.001)
 
         table_path = os.path.join(dirname,  'modules', 'table.urdf')
-        p.loadURDF(table_path, [-0.6, -4.6, 0.95], p.getQuaternionFromEuler([0, 0, np.pi/2]),
+        p.loadURDF(table_path, [-1.2, -0.6, 0.95], p.getQuaternionFromEuler([0, 0, np.pi/2]),
                    useFixedBase=True, globalScaling=0.001)
 
         #load Robots__________________________________________________
@@ -134,11 +140,14 @@ class LaserTrackerEnv:
                 'q6': -0.09411935083551597
             }
             self.robot.set_joint_position(joint_angles)
+            self.second_robot.set_joint_position(joint_angles)
             p.stepSimulation()
 
 
-        for position,orientation,_ in self.path:
[…]

### 2024-07-25 code/transformer_cell/main.py

@@ -57,7 +57,7 @@ if __name__ == '__main__':
     p.loadURDF(dmg_path, [0,0,0], p.getQuaternionFromEuler([math.pi/2,0,0]), useFixedBase=True, globalScaling=0.001)
     mm_path = os.path.join(dirname,
                              'Objects', 'modules', 'milling_module.urdf')
-    p.loadURDF(mm_path, [-0.6,-4.6,0.95], p.getQuaternionFromEuler([0,0,math.pi/2]), useFixedBase=True, globalScaling=0.001)
+    p.loadURDF(mm_path, [-3,-2.6,0.95], p.getQuaternionFromEuler([0,0,math.pi/2]), useFixedBase=True, globalScaling=0.001)
     table_path = os.path.join(dirname,
                            'Objects', 'modules', 'table.urdf')
     p.loadURDF(table_path, [-0.6, -4.6, 0.95], p.getQuaternionFromEuler([0, 0, math.pi/2]), useFixedBase=True,

### 2024-07-25 code/optical_marker.py
added distance to obective and fixed module collision shape

@@ -84,6 +84,21 @@ class ActiveMarker(pi.EndeffectorTool):
 
         return visibility
 
+    def compute_distance(self):
+        """computes the distance betwen the marker and the tracker
+
+        Returns:
+            list: A list which contains the distance of each marker for its corresponding tracker
+        """
+        num_markers = len(self.marker_positions)
+        distances = np.zeros(num_markers)
+        current_marker_positions, current_marker_orientations = self.get_marker_poses_orientations()
+        for i in range(num_markers):
+            ray_start_pos = self.tracker_positions[i]
+            ray_end_pos = current_marker_positions[i]
+            distances[i] = np.linalg.norm(np.array(ray_start_pos) - np.array(ray_end_pos))
+        return distances
+
     def get_marker_poses_orientations(self):
         """ Returns the positions and orientations of the markers in the world frame.
 

### 2024-07-25 code/paper_experiment.py

@@ -124,6 +124,7 @@ class LaserTrackerEnv:
             and marker poses. It then returns the cost for each timestep and particle.
         """
         visibility_index = []
+        distance_index = []
         self.reset()
 
         self.marker.set_optical_system_parameters(particles.T)
@@ -151,11 +152,14 @@ class LaserTrackerEnv:
             for _ in range(100):
                 p.stepSimulation()
             visibility_index.append(self.marker.compute_visibility( ))
+            distance_index.append(self.marker.compute_distance())
 
         visibility_index = np.array(visibility_index)
+        distance_index = np.array(distance_index)
 
         visibility_index = visibility_index.T
-        return visibility_index
+        distance_index = distance_index
+        return visibility_index, distance_index
 
     def set_timestep_length(self, time_step_length):
         """Sets the timestep between subsequent step function calls.
@@ -181,12 +185,19 @@ def run_experiment():
     console = Console()
     def objective_function(particles):
 
-        visibility_cost = env.run_simulation(particles)
+        visibility_cost,distance_cost = env.run_simulation(particles)
         visibility_cost_sum = -1*np.sum(visibility_cost,axis=1)
 
-        #TODO add cost that keeps the tracker as close as possible to the surface of the robot
+        # cost that keeps the tracker as close as possible to the surface of the robot
         marker_positions = particles[3:6,:]
         deviation_cost = np.linalg.norm(marker_positions, axis=0)
+
+        # cost that keeps the distance cost as low as possible so long as it is not below 1.2
+        distance_cost = np.sum(distance_cost,axis=0)/len(distance_cost)
+
+        distance_cost = np.array([1e2 if x < 1.2 else x for x in distance_cost])
+
+
 
         #constraint costs
 
@@ -208,7 +219,7 @@ def run_experiment():
 
 
 
-        costs = visibility_cost_sum+ deviation_cost+height_constraint + radius_constraint + height_constraint
+        costs = visibility_cost_sum+ +distance_cost+ deviation_cost+height_constraint + radius_constraint + height_constraint
 
         print(costs)
 
@@ -220,7 +231,7 @@ def run_experiment():
     n_particles = 100
 
 
-    initial_tracker_positions = np.random.rand(3, n_particles)*2
+    initial_tracker_positions = np.random.rand(3, n_particles)*3
[…]

### 2024-07-25 paper/paper.tex
updaeted penalty for minimization and range objective in paper

@@
     f(p_m,p_l) = \begin{cases}
-    1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
+    -1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
     0 & \text{otherwise}
@@
     g(p_m,p_l) = \begin{cases}
-    1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
+    -1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
     0 & \text{otherwise}
@@
 \begin{equation}
-    g(p_m,p_l) = \cos(\alpha)
+    g(p_m,p_l) = -\cos(\alpha)
 \end{equation}
@@
 \begin{equation}
-    h(p_m,p_l) = \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l)
+    h(p_m,p_l) = \int \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l) dt
     \label{eq:objective}
@@
 where $\lambda$ is a weighting factor between 0 and 1 that can be adjusted to prioritize one component over the other.
+The intergral is taken over the entire trajectory of the manufacturing system that needs to be observed by the laser tracker.
 After defining the objective function we can now define the design variables and constraints.
@@
 The laser tracker also has a limited range of measurement.
-Here we can define a constraint that the distance between the laser tracker and the marker is within the range of the laser tracker:
+Here we could define a constraint that the marker needs to be within the measurement range of the laser tracker.
+But in practice the laser tracker works better the closer the marker is to the laser tracker so long it is in the minimum range.
+This can be encoded using a penalty function:
 \begin{equation}
-    d_{min} \leq ||p_{p_m} - p_{p_l}|| \leq d_{max}
+    p(p_m,p_l) =  \begin{cases}
+        d & \text{if } \text{marker is in measurement range of the laser tracker} \\
+        \rho & \text{otherwise}
+    \end{cases}
 \end{equation}
-where $d_{min}$ and $d_{max}$ are the minimum and maximum range of the laser tracker.
-The final design space is then 9 dimensional, 6 for the marker pose and 3 for the laser tracker pose.
+where $d$ is the distance of the marker to the laser tracker and $\rho$ is a large penalty factor.
+The final optimization problem can be formulated as follows:
+\begin{equation}
+    \begin{split}
+    \max_{p_m,p_l} & h(p_m,p_l) \\
+    \text{s.t.} & \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
+    & 0 \leq p_{p_m,z} \leq h \\
+    & h_{min} \leq p_{p_l,z} \leq h_{max} \\
+    & p(p_m,p_l) \geq 0
+    \end{split}
+\end{equation}
 
@@
         0 & \text{if } x \text{ satisfies all constraints} \\
-        -\rho & \text{otherwise}
+        \rho & \text{otherwise}
     \end{cases}

### 2024-07-25 code/paper_experiment.py
fixed bug in distance constraint where 1.2 was evaluated after sum

@@ -193,9 +193,13 @@ def run_experiment():
         deviation_cost = np.linalg.norm(marker_positions, axis=0)
 
         # cost that keeps the distance cost as low as possible so long as it is not below 1.2
-        distance_cost = np.sum(distance_cost,axis=0)/len(distance_cost)
 
-        distance_cost = np.array([1e2 if x < 1.2 else x for x in distance_cost])
+        # go through all the values in the distance cost and replace them with 1e2 if they are below 1.2
+        # bore summing over the trajectory
+        distance_cost = np.where(distance_cost < 1.2, 1e2, distance_cost)
+
+        distance_cost = np.sum(distance_cost,axis=0)
+
 
 
 
@@ -216,10 +220,7 @@ def run_experiment():
 
 
 
-
-
-
-        costs = visibility_cost_sum+ +distance_cost+ deviation_cost+height_constraint + radius_constraint + height_constraint
+        costs = visibility_cost_sum+ distance_cost + deviation_cost+height_constraint + radius_constraint + height_constraint
 
         print(costs)
 
@@ -231,7 +232,7 @@ def run_experiment():
     n_particles = 100
 
 
-    initial_tracker_positions = np.random.rand(3, n_particles)*3
+    initial_tracker_positions = np.random.rand(3, n_particles)*10
     initial_marker_positions = np.random.rand(3, n_particles)*0.6
     initial_marker_orientations = np.random.rand(3, n_particles)*2*np.pi
 

### 2024-07-25 code/paper_evaluation.py
changed visualisation to show all similarly good results

@@ -24,7 +24,7 @@ plt.figure(figsize=(3, 2))
 sns.lineplot(x=range(objective_log.shape[0]), y=mean_values, color=kit_deep_blue, label='Mean')
 
 # Plot the hull curve
-plt.fill_between(range(objective_log.shape[0]), lower_boundary, upper_boundary, color=see_through_blue, label='5th-95th percentile range')
+plt.fill_between(range(objective_log.shape[0]), lower_boundary, upper_boundary, color=see_through_blue, label='particle range')
 
 plt.xlabel("Solver iteration")
 plt.ylabel("Objective function")

### 2024-07-25 code/paper_visualization.py

@@ -1,8 +1,58 @@
+import numpy as np
+import os
+import tkinter as tk
+from tkinter import ttk
 from paper_experiment import LaserTrackerEnv
 import pybullet as p
 import pybullet_industrial as pi
-import numpy as np
-import os
+
+class ParticleSelectorApp:
+    def __init__(self, master, particles, objective_log, env, threshold=0.01):
+        self.master = master
+        self.particles = particles
+        print(self.particles)
+        self.objective_log = objective_log
+        print(self.objective_log)
+        self.env = env
+
+        # Find the best objective value
+        min_objective_value = np.min(objective_log[-1])
+
+        # Filter particles within the threshold
+        self.filtered_indices = np.where(objective_log[-1] <= min_objective_value * (1 + threshold))[0]
+        self.filtered_particles = particles[:, self.filtered_indices]
+
+        self.master.title("Select Particle Solution")
+
+        self.label = tk.Label(master, text="Select a Particle Solution:")
+        self.label.pack()
+
+        self.combobox = ttk.Combobox(master)
+        self.combobox.pack()
+        self.combobox['values'] = [f"Solution {i}" for i in self.filtered_indices]
+        self.combobox.bind("<<ComboboxSelected>>", self.update_simulation)
+
+        laser_tracker_path = os.path.join(os.path.dirname(__file__), 'transformer_cell', 'Objects', 'laser_tracker.urdf')
+        self.laser_tracker = p.loadURDF(laser_tracker_path, [0,0,0], p.getQuaternionFromEuler([0, 0, 0]), useFixedBase=True, globalScaling=0.001)
+
+        self.selected_particle = None
+
+    def update_simulation(self, event):
+        selected_index = self.combobox.current()
+        best_particle = self.particles[:, selected_index]
+
+        # reshape to 9x1
+        best_particle = best_particle.reshape(9, 1)
+
+        laser_tracker_position = best_particle[:3].flatten()
+
+        # move the laser tracker to the selected position
+        p.resetBasePositionAndOrientation(self.laser_tracker, laser_tracker_position, [0, 0, 0, 1])
+
+
+        self.selected_particle = best_particle
+        self.env.run_simulation(best_particle)
+
 
 if __name__ == "__main__":
     particle_log = np.load("particle_log.npy")
@@ -11,29 +61,6 @@ if __name__ == "__main__":
 
     last_particle_round = particle_log[-1]
 
[…]

### 2024-07-25 code/paper_visualization.py
tried different iteration counts for solutions

@@ -7,7 +7,7 @@ import pybullet as p
 import pybullet_industrial as pi
 
 class ParticleSelectorApp:
-    def __init__(self, master, particles, objective_log, env, threshold=0.01):
+    def __init__(self, master, particles, objective_log, env, threshold=1):
         self.master = master
         self.particles = particles
         print(self.particles)
@@ -16,10 +16,10 @@ class ParticleSelectorApp:
         self.env = env
 
         # Find the best objective value
-        min_objective_value = np.min(objective_log[-1])
+        min_objective_value = np.min(objective_log)
 
         # Filter particles within the threshold
-        self.filtered_indices = np.where(objective_log[-1] <= min_objective_value * (1 + threshold))[0]
+        self.filtered_indices = np.where(objective_log<= min_objective_value * (1 + threshold))[0]
         self.filtered_particles = particles[:, self.filtered_indices]
 
         self.master.title("Select Particle Solution")
@@ -59,8 +59,9 @@ if __name__ == "__main__":
     objective_log = np.load("objective_log.npy")
     env = LaserTrackerEnv(rendering=True)
 
-    last_particle_round = particle_log[-1]
+    last_particle_round = particle_log[70]
+    last_objective_round = objective_log[70]
 
     root = tk.Tk()
-    app = ParticleSelectorApp(root, last_particle_round, objective_log, env)
+    app = ParticleSelectorApp(root, last_particle_round, last_objective_round, env)
     root.mainloop()

### 2024-07-29 paper/paper.tex
added marker vis and rephrased experiment section

@@
 
-\author{Jan Baumgärtner$^{1}$,  Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
+\author{Jan Baumgärtner$^{1}$, Max Goebels,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
 \thanks{*The authors would like to thank the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg for the financial support of the project within the InnovationsCampus Future Mobility.
 }% <-this % stops a space
-\thanks{$^{1}$Jan Baumgärtner, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
+\thanks{$^{1}$Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
         Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
-        {\tt\small jan.baumgaertner@kit.edu}}%
+        {\tt\small max.goebels@kit.edu}}%
 }
@@
 If we still want the marker close to the surface of the system such as when mounting them magnetically we can add a penalty function to the objective function that penalizes markers that are too far away from the flange.
-This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.
-
+This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.\\
+\\
 The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
@@
                 \hline
-                Number of particles & 100 \\
-                Number of iterations & 1000 \\
-                Inertia weight & 0.9 \\
-                Cognitive coefficient & 2.0 \\
-                Social coefficient & 2.0 \\
+                Number of particles & 500 \\
+                Number of iterations & 100 \\
+                Inertia weight & 0.8 \\
+                Cognitive coefficient & 0.1 \\
+                Social coefficient & 0.1 \\
         \end{tabular}
@@
 \section{Experiment Setup}
-To test the performance of our optimiser, we need to define a set of test cases.
-These consist of different trajectories of the two robots shown in figure \ref{fig:simulation}.
-Both robots are Comau NJ290 industrial robots with a reach of 2.9 metres.
-The right robot is used to service the CNC machine and the additive manufacturing system, while the left robot is used to perform a high-precision manufacturing task.
-To properly control this task, the left robot needs to be tracked by the laser tracker.
-We test whether our algorithm is able to find a placement of the laser tracking system and markers that can track the robot over its entire trajectory.
-We compare this to the performance of the original population of marker and laser tracker placements used to start the optimisation.
+To test the performance of our optimizer, we need to define a testcase.
+Both robots are Comau NJ290 industrial robots with a reach of 2.9 meters.
+The right robot services the CNC machine and the additive manufacturing system, while the left robot performs a high-precision manufacturing task.
+For this test, the left robot follows a rectangular trajectory with dimensions of 0.4 meters in length and 0.6 meters in width, as depicted in Figure \ref{fig:trajectory}.
+The aim is now to find a laser tracker and marker placement that can track the left robot throughout its entire trajectory.
+We compare this to the performance of the initial population of marker and laser tracker placements used to start the optimization process.
+Figure \ref{fig:iterations} illustrates the performance of the algorithm over multiple iterations.
+The results show that the algorithm converges to a solution in roughly 40 iterations, which is relatively fast for an optimization approach.
+This speed is likely due to the existence of numerous optimal placements for the laser tracker and markers.
+The final solution is also depicted in Figure \ref{fig:trajectory}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/iterations.pdf}
+\caption{Performance of the algorithm over the iterations.}
+\label{fig:iterations}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.8\textwidth]{figures/trajectory.png}
+\caption{Trajectory of the robot and final laser tracker and marker placements.}
+\label{fig:trajectory}
+\end{figure}
 

### 2024-07-29 code/optical_marker.py

@@ -132,6 +132,11 @@ class ActiveMarker(pi.EndeffectorTool):
         self.marker_positions = [particle[3:6] for particle in particles]
         self.marker_orientations = [p.getQuaternionFromEuler(particle[6:9]) for particle in particles]
 
+        # draw markers
+        for i in range(len(self.tracker_positions)):
+            pi.draw_coordinate_system(self.marker_positions[i], self.marker_orientations[i],parent_id = self.urdf)
+
+
 class TestActiveMarker(unittest.TestCase):
 
     def setUp(self):

### 2024-07-29 code/paper_evaluation.py

@@ -26,6 +26,9 @@ sns.lineplot(x=range(objective_log.shape[0]), y=mean_values, color=kit_deep_blue
 # Plot the hull curve
 plt.fill_between(range(objective_log.shape[0]), lower_boundary, upper_boundary, color=see_through_blue, label='particle range')
 
+# y axis log scale
+plt.yscale("log")
+
 plt.xlabel("Solver iteration")
 plt.ylabel("Objective function")
 plt.legend()

### 2024-07-29 code/paper_experiment.py

@@ -229,10 +229,12 @@ def run_experiment():
 
     env = LaserTrackerEnv()
 
-    n_particles = 100
+    n_particles = 500
 
 
-    initial_tracker_positions = np.random.rand(3, n_particles)*10
+    initial_tracker_positions = np.random.rand(3, n_particles)*3
+    initial_tracker_positions[0] += -3.5
+    initial_tracker_positions[1] += -3.5
     initial_marker_positions = np.random.rand(3, n_particles)*0.6
     initial_marker_orientations = np.random.rand(3, n_particles)*2*np.pi
 

### 2024-07-29 code/paper_visualization.py

@@ -59,8 +59,8 @@ if __name__ == "__main__":
     objective_log = np.load("objective_log.npy")
     env = LaserTrackerEnv(rendering=True)
 
-    last_particle_round = particle_log[70]
-    last_objective_round = objective_log[70]
+    last_particle_round = particle_log[-1]
+    last_objective_round = objective_log[-1]
 
     root = tk.Tk()
     app = ParticleSelectorApp(root, last_particle_round, last_objective_round, env)

### 2024-07-29 code/paper_experiment.py
added concave mesh to experiment

@@ -22,7 +22,7 @@ class LaserTrackerEnv:
             self.physics_client = p.connect(p.DIRECT)
         else:
             self.physics_client = p.connect(p.GUI, options='--background_color_red=1 ' +
-                                            '--background_color_green=1 ' +
+                                            '--background_color_green=0 ' +
                                             '--background_color_blue=1')
             p.configureDebugVisualizer(p.COV_ENABLE_GUI, 0)
 
@@ -33,7 +33,7 @@ class LaserTrackerEnv:
         p.setTimeStep(self._time_step_length)
 
         if path is None:
-            target_position = np.array([-2.4,-3,0.95])+np.array([-0.6,0.45,0.4])
+            target_position = np.array([-2.4,-3,1.05])+np.array([-0.6,0.45,0.4])
             steps = 30
             self.path = pi.build_box_path(
                 target_position, [0.5, 0.6], 0.1, [0, 0, 0, 1], steps)
@@ -72,7 +72,7 @@ class LaserTrackerEnv:
 
         mm_path = os.path.join(dirname,  'modules', 'milling_module.urdf')
         p.loadURDF(mm_path, [-2.4,-3,0.95], p.getQuaternionFromEuler([0,0,np.pi/2]),
-                   useFixedBase=True, globalScaling=0.001)
+                   useFixedBase=True, globalScaling=0.001,flags= p.GEOM_FORCE_CONCAVE_TRIMESH)
 
         table_path = os.path.join(dirname,  'modules', 'table.urdf')
         p.loadURDF(table_path, [-1.2, -0.6, 0.95], p.getQuaternionFromEuler([0, 0, np.pi/2]),
@@ -142,6 +142,13 @@ class LaserTrackerEnv:
             }
             self.robot.set_joint_position(joint_angles)
             self.second_robot.set_joint_position(joint_angles)
+            for _ in range(50):
+                p.stepSimulation()
+
+
+        self.marker.set_tool_pose(self.path.positions[:,0]+np.array([0,0,1.3]),self.path.orientations[:,0])
+        pi.draw_coordinate_system(self.path.positions[:,0]+np.array([0,0,1.3]),self.path.orientations[:,0])
+        for _ in range(100):
             p.stepSimulation()
 
 
@@ -264,10 +271,10 @@ def run_experiment():
 
 
[…]

### 2024-07-29 paper/paper.tex
added citations

@@
 \usepackage{makecell}
+\usepackage{biblatex}
+\bibliography{bibliography}
 
@@
 
-\author{Jan Baumgärtner$^{1}$, Max Goebels,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
-\thanks{*The authors would like to thank the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg for the financial support of the project within the InnovationsCampus Future Mobility.
-}% <-this % stops a space
+\author{Jan Baumgärtner$^{1*}$, Max Goebels$^{1*}$,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
+\thanks{*These authors contributed equally to this work.}% <-this % stops a space
 \thanks{$^{1}$Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
@@
 Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
-Similar to approaches such as \cite{ieee_sensors} we divide the visibility into two components.
+Similar to approaches such as~\cite{ieee_sensors} we divide the visibility into two components.
 Line of sight visibility and field of view visibility.
-Both  are depicted in figure \ref{fig:visibility}.
+Both  are depicted in figure~\ref{fig:visibility}.
 \begin{figure}
@@
 \end{equation}
+Apart from line of sight we also want to optimize the measurment performance of the laser tracker.
+The accuracy of the lasertracker is mainly dependent on the distance between the marker and the laser tracker.
+The further away the marker is the less accurate the measurment will be. However there is also a minimum measurment distance that the laser tracker can handle.
+To encode this we can define a penalty function that penalizes markers that are too far away from the laser tracker.
+\begin{equation}
+    p(p_m,p_l) =  \begin{cases}
+        d & \text{if marker is in measurement range of the laser tracker} \\
+        \rho & \text{otherwise}
+    \end{cases}
+\end{equation}
+where $d$ is the distance between the marker and the laser tracker, and $\rho$ is a large penalty factor.
 The final objective function is then a weighted sum of the two components:
 \begin{equation}
-    h(p_m,p_l) = \int \lambda f(p_m,p_l) + (1-\lambda) g(p_m,p_l) dt
+    h(p_m,p_l) = \int f(p_m,p_l) +  g(p_m,p_l) + p(p_m,p_l) \, dt
     \label{eq:objective}
 \end{equation}
-where $\lambda$ is a weighting factor between 0 and 1 that can be adjusted to prioritize one component over the other.
+
 The intergral is taken over the entire trajectory of the manufacturing system that needs to be observed by the laser tracker.
-After defining the objective function we can now define the design variables and constraints.
-The design variables are the pose of the laser tracker $p_l$ and the pose of the marker $p_m$.
-However the pose of these objects is not arbitrary.
-For the purposes of this paper we assume that the marker is attached to the flange of the manufacturing system.
-This means that we only define a Transformation matrix $T_{m}$ that describes the pose of the marker relative to the flange.
-The final pose of the marker $p_m$ is then given by the flange transformation $T_{f}(q)$ where $q$ is the joint configuration of the robot as well as the transformation $T_{m}$:
+Having defined the objective function, we can proceed to define the design variables and constraints.
+The design variables are the pose of the laser tracker ($p_l$) and the pose of the marker ($p_m$).\\
+\\
+First, we assume the marker is attached to the flange of the manufacturing system, meaning we only need to define a transformation matrix $T_{m}$ that describes the marker's pose relative to the flange.
+The final pose of the marker ($p_m$) is then given by combining the flange transformation $T_{f}(q)$, where $q$ is the robot's joint configuration, with the transformation $T_{m}$:
+
 \begin{equation}
@@
 \end{equation}
-While the matrix $T_m$ is a 4x4 matrix, we can parameterize it using the position of the marker $p_{p_m}$ and the orientation of the marker $p_{o_m}$.
-This results in a 6 dimensional design space for the marker pose.
-This space is further constraint to avoid any unrealistic marker positions (such as 10 meters above the robot).
-Ideally the marer should be attached close to the surface of the manufacutring system to avoid any unneccessarily large constructions that might collide with the environment or cause vibrations.
-This can be encoded using a cylindrical constraint on the position of the marker:
+
+While $T_{m}$ is a 4x4 matrix, it can be parameterized using the marker's position ($p_{p_m}$) and orientation ($p_{o_m}$), resulting in a 6-dimensional design space for the marker pose.
+To avoid unrealistic positions, we constrain this space to keep the marker close to the manufacturing system's surface.
+This constraint can be represented cylindrically:
 \begin{equation}
@@
     \end{split}
-\end{equation}
-where $r$ is the radius of the cylinder and $h$ is the height of the cylinder.
-The marker of course also cant be within the mesh of the manufacturing system, but since this would result in a line of sight visibility of 0 we do not need to encode this constraint explicitly.
-If we still want the marker close to the surface of the system such as when mounting them magnetically we can add a penalty function to the objective function that penalizes markers that are too far away from the flange.
-This will naturally force the optimizer to place the marker close to the surface of the manufacturing system.\\
+\end{equation}I
+
+where $r$ is the cylinder's radius and $h$ is its height.
+Additionally, to ensure the marker remains near the surface, a penalty function can be added to the objective function to penalize markers positioned too far from the flange.\\
 \\
-The pose of the laser tracker is in reality also defined by a transformation matrix $T_{l}$.
-However since the laser tracker has a spherical measurment volume we only need to define the position of the laser tracker $p_{p_l}$, which is a three dimensional vector.
-The laser tracker is also mounted on a stand, in principle this can vary in height but only in a limited range.
-This can be encoded using a simple constraint:
+The laser tracker's pose is defined by a transformation matrix $T_{l}$, but due to its spherical measurement volume, only its position ($p_{p_l}$) needs to be specified.
+The laser tracker is mounted on a stand with a height range constraint:
 \begin{equation}
@@
 \end{equation}
-where $h_{min}$ and $h_{max}$ are the minimum and maximum height of the laser tracker stand.
-The laser tracker also has a limited range of measurement.
-Here we could define a constraint that the marker needs to be within the measurement range of the laser tracker.
-But in practice the laser tracker works better the closer the marker is to the laser tracker so long it is in the minimum range.
-This can be encoded using a penalty function:
-\begin{equation}
-    p(p_m,p_l) =  \begin{cases}
-        d & \text{if } \text{marker is in measurement range of the laser tracker} \\
-        \rho & \text{otherwise}
-    \end{cases}
-\end{equation}
-where $d$ is the distance of the marker to the laser tracker and $\rho$ is a large penalty factor.
-The final optimization problem can be formulated as follows:
+Where $h_{min}$ and $h_{max}$ are the minimum and maximum heights of the stand, respectively.
+Using the design variables and constraints of the marker and laser tracker, the optimization problem can be formulated as follows:
 \begin{equation}
     \begin{split}
-    \max_{p_m,p_l} & h(p_m,p_l) \\
-    \text{s.t.} & \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
-    & 0 \leq p_{p_m,z} \leq h \\
-    & h_{min} \leq p_{p_l,z} \leq h_{max} \\
-    & p(p_m,p_l) \geq 0
+        \max_{p_m,p_l} & \ h(p_m,p_l) \\
+        \text{s.t.} & \ \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} \leq r \\
+        & \ 0 \leq p_{p_m,z} \leq h \\
+        & \ h_{min} \leq p_{p_l,z} \leq h_{max} \\
     \end{split}
@@
 
+\section{Acknowledgements}
+
+%Bibliography
+%TODO checken ob das stimmt in template
+\printbibliography
+
 \end{document}

### 2024-07-29 paper/paper.tex
added new results with correct collision mesh

@@
 \pagestyle{empty}
+
+\section{Introduction}
+
+\section{State of the art}
 
@@
 On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
-These two properties make the problem well suited for the particle swarm optimization (pso) algorithm.
-Particle swarm optimization (PSO) is a computational method for finding the optimal solution to a problem
+These two properties make the problem well suited for the particle swarm optimization (PSO) algorithm.
+Particle swarm optimization is a computational method for finding the optimal solution to a problem
  It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
@@
 In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
-The best particle is the one that maximizes the objective function defined in equation \ref{eq:objective}.
-The hyperparameters we need to tune are collected in table \ref{tab:hyperparameters}.
+The best particle is the one that maximizes the objective function defined in equation~\ref{eq:objective}.
+The hyperparameters we need to tune are collected in table~\ref{tab:hyperparameters}.
 Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
@@
 
-\section{Experiment Setup}
+\section{Sample Testcase}
 To test the performance of our optimizer, we need to define a testcase.

### 2024-07-29 code/paper_experiment.py

@@ -22,7 +22,7 @@ class LaserTrackerEnv:
             self.physics_client = p.connect(p.DIRECT)
         else:
             self.physics_client = p.connect(p.GUI, options='--background_color_red=1 ' +
-                                            '--background_color_green=0 ' +
+                                            '--background_color_green=1 ' +
                                             '--background_color_blue=1')
             p.configureDebugVisualizer(p.COV_ENABLE_GUI, 0)
 
@@ -271,10 +271,10 @@ def run_experiment():
 
 
 if __name__ == "__main__":
-    #run_experiment()
-    env = LaserTrackerEnv(rendering=True)
+    run_experiment()
+    #env = LaserTrackerEnv(rendering=True)
 
-    while True:
-        env.run_simulation(np.array([[-4,-1,3,0,0,0.3,0,0,0],[-3,-5,1.1,0,-0.3,0.1,0,0,0]]).T)
+    #while True:
+    #    env.run_simulation(np.array([[-4,-1,3,0,0,0.3,0,0,0],[-3,-5,1.1,0,-0.3,0.1,0,0,0]]).T)
 
 

### 2024-08-04 paper/paper.tex
added trajectory graphic to paper
added first pass at state of the art and abstract

@@
 
+\section*{Abstract}
+In recent years, modular and reconfigurable manufacturing systems have gained attention for their adaptability. 
+While system design is now highly automated, designing the verification setup remains a manual design problem involving tasks like placing external sensors to measure performance. 
+This paper argues that design optimization approaches for manufacturing systems can also automate the design of the verification setup.
+This is demonstrated using the placement of a laser tracker system with markers that due to their line-of-sight complexity and simultaneous measurement of multiple systems can be challenging to design.
+We present a straightforward optimization approach to demonstrate potential automation in this area.
 \section{Introduction}
@@
 \section{State of the art}
+The work presented in this paper lies at the intersection of two research fields: optimization of manufacturing systems and sensor placement optimization.
+The question this work first has to answer is what do we actually verify in a manufacturing system.
+The first thing we have to clarify here is that we are not looking at product based verification but system based verification.
+This means we dont want to look at the resulting quality of the product but at the performance of the manufacturing system itself.
+The main reason for this is tha the reconfigurable manufacturing systems we are looking at are not varying in processes but in how the machine achieves the same process.
+This means we dont want to measure the process influence but the machine influence.
+In the framework of \cite{reconfigurable_production} this means we are mainly dealing with the reconfiguration design issue.
+Examples of such systems include the robot cell optimization problem proposed in \cite{previous_work} or \cite{stiffness_placement} 
+but also robot design optimization problems such as \cite{task_synthesis}, \cite{ad-hoc_manipulator} or\cite{multi_objective}.
+These works are very robotics specific for a more manufacturing based perspective we can look at works such as \cite{johannes_1} or \cite{johannes_2}.
+The main considerations in these works where the possible payload, accuracy, stiffness, force, weight and possible motion of the systems.
+The important point is that in order to verify that the system configuration that was optimized is indeed capable of achieving the desired performance we need to measure the performance of the system.
+Since most configurable manufacturing systems are not capable of directly measuring their accuracy, force or stiffness we need to use external sensors to measure these values.
+This requires a verification setup that is capable of measuring the performance of the manufacturing system.
+If we want to automate this process we are dealing with the domain of sensor placement optimization.
+The main goal of sensor placement optimization is to find the optimal placement of sensors in a system such that the performance of the system can be measured as accurately as possible.
+This is a well studied field in the context of structural health monitoring where a comprehensive overview can be found in \cite{shms}.
+Important algorithms in this field are the genetic algorithm, particle swarm optimization and simulated annealing \cite{shms}.
+Borrowing from this field we will also use a particle swarm optimization algorithm to solve our optimization problem.
+The paper also explores a host of different measurement types such as Velocity, displacement, acceleration, strain, temperature or pressure.
+Howeverwhile it is possible to borrow from this field additional challenges are introduces by the fact that manufacutring systems are typically moving
+and that the sensor placement thus has to take into account the total trajectory of the manufacturing system.
+Only recently has this challenge been started to be addressed in the context of manufacturing systems \cite{ieee_sensors}.
+Allhough this early work focuses on process control and not system verification its methodologies can be applied to our problem.
+Mainly that instead of optimizing a single objective function we need to optimize a objective integrated over the total process time.
+However since we are not looking at cameras such as the previous paper we need to derive our own visibility model reflecting the measurement characteristics of a laser tracker.
 
@@
 There are a large number of possible optimization algorithms that can be used to solve this problem.
-Following the no free lunch theorem we want to pick the algorithm that makes the most use of the underlying structure of the problem. %TODO citation needed
+Following the no free lunch theorem \cite{no_free_lunch_theorem} we want to pick the algorithm that makes the most use of the underlying structure of the problem.
 In this case the problem is a continuous optimization problem with a relatively low number of design variables.
 On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
-These two properties make the problem well suited for the particle swarm optimization (PSO) algorithm.
-Particle swarm optimization is a computational method for finding the optimal solution to a problem
+These two properties make the problem well suited for the particle swarm optimization (PSO) algorithm \cite{pso}.
+Particle swarm optimization is a computational method for finding the optimal solution to a problem.
 It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
@@
 \centering
-\includegraphics[width=0.8\textwidth]{figures/trajectory.png}
+\includegraphics[width=\textwidth]{figures/trajectory.png}
 \caption{Trajectory of the robot and final laser tracker and marker placements.}

### 2024-08-04 paper/paper.tex
polished state of the art

--- 2024-07-29 paper/paper.tex
@@
 \section{State of the art}
+The work presented in this paper intersects two research fields: manufacturing systems optimization and sensor placement optimization. 
+Our primary question is what aspects of a manufacturing system we need to verify. 
+We focus on system-based verification rather than product-based, aiming to measure the performance of the manufacturing system, not the resulting product quality.
+In particular we focuss on approaches that solves what~\cite{reconfigurable_production} calls reconfiguration design problem.
+This involves either combining fixed modules to form a system or directly optimizing the sytems physical design.
+ Examples include robot cell optimization problems~\cite{previous_work, stiffness_placement} and robot design optimization problems~\cite{task_synthesis, ad-hoc_manipulator, multi_objective}. 
+ For a manufacturing perspective, works such as~\cite{johannes_1, johannes_2} are relevant.
+ All of these approaches focus on phyiscal capabilities such as payload, accuracy, stiffness, force, weight, and motion.
+
+To verify that the optimized system configuration achieves the desired performance, we need to measure the system's performance using external sensors, 
+however in all of the above works the sensor placement for verification (if done at all) is done manually.
+This can be a challening task which brings us to the field of sensor placement optimization.
+The goal here is to find the optimal sensor placement for accurate system performance measurement.
+A adjecent engineering discipline in which sensor placement optimization is a already well studied is structural health monitoring~\cite{shms}.
+Here key algorithms where already identified such as genetic algorithms, particle swarm optimization, and simulated annealing~\cite{shms}.
+Based on these works we will also use a particle swarm optimization algorithm for our problem.
+However the field of structural health monitoring mainly considers static systems while our manufacturing systems are by nature dynamic.
+This introduces additional challenges since the sensor placement needs to account for the systems total trajectory.
+This challenge has only recently been addressed in manufacturing systems~\cite{ieee_sensors}.
+Although this early work focuses on process control, not system verification, its methodologies are applicable,
+emphasizing optimizing an objective integrated over total process time. 
+However, since we are not using cameras as in the previous study, we must derive our own visibility model reflecting the measurement characteristics of a laser tracker.
+This will be the focus of the next section.
 

### 2024-08-04 paper/paper.tex
added conclusion

@@
 The final solution is also depicted in Figure \ref{fig:trajectory}.
+In conclusion the results show that the optimization approach is capable of finding a suitable laser tracker and marker placement for the manufacturing system.
+We have also verified that the algorithm is capable of handling the dynamic nature of the manufacturing system, as the optimization is performed over the entire trajectory of the robot.
 
@@
 
+\section{Conclusion}
+
+In this paper, we argue that automating the verification setup in manufacturing systems is feasible using optimization methods.
+We demonstrated this with a particle swarm optimization algorithm that optimizes the placement of a laser tracker and markers. 
+Although initially labor-intensive, this approach proves beneficial for frequently reconfiguring systems as it requires only a simulation for each verification. 
+Our results confirm the algorithm's effectiveness in determining optimal sensor placements, significantly reducing the manual effort required by engineers. 
+We believe sensor placement optimization is a promising tool for researchers to verify system performance. 
+However, the field is nascent with several unanswered questions about integrating optimization in reconfiguration processes, 
+the potential for setup reuse across different designs, and developing multiple optimal solutions. 
+We hope this paper serves as a catalyst for further research in this domain.
+
 \section{Acknowledgements}

### 2024-08-05 paper/paper.tex
added intro, some templating.

@@
 
-%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper
-
-\IEEEoverridecommandlockouts                              % This command is only needed if
-                                                          % you want to use the \thanks command
-
-\overrideIEEEmargins                                      % Needed to meet printer requirements.
-
-%In case you encounter the following error:
-%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
-%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
-%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
-%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
-%\pdfobjcompresslevel=0
-%\pdfminorversion=4
-
-% See the \addtolength command later in the file to balance the column lengths
-% on the last page of the document
+% to typeset URLs, URIs, and DOIs
+\usepackage{url}
+\def\UrlFont{\rmfamily}
 
@@
 
-\title{\LARGE \bf
-Laser Tracker Placement Optimization for highly flexible manufacturing systems
+\begin{document}
+\mainmatter              % start of a contribution
+\title{
+Laser tracker placement optimization for highly flexible manufacturing systems
+}
+\titlerunning{Laser tracker placement}
+
+% latex magic to format multiple authors affiliations
+\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
+\makeatletter
+% *, 2, 3, ...
+\renewcommand*{\@fnsymbol}[1]{\ifcase#1\or\@arabic{#1}\else*\fi}
+\makeatother
+
+\author{Jan Baumgärtner \inst{1,2}, 
+Max Goebels \inst{1,2},   Alexander Puchta \inst{1,2} and Jürgen Fleischer\inst{1}% <-this % stops a space
+}
+\institute{wbk Institute of Production Science, Karlsruhe Institute of Technology, 
+76131 Karlsruhe, Germany \email{max.goebels@kit.edu}
+\and
+These authors contributed equally to this work.
 }
 
-\author{Jan Baumgärtner$^{1*}$, Max Goebels$^{1*}$,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
-\thanks{*These authors contributed equally to this work.}% <-this % stops a space
-\thanks{$^{1}$Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
-        Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
-        {\tt\small max.goebels@kit.edu}}%
-}
-
-\begin{document}
+\authorrunning{Baumgärtner et al.}
+\tocauthor{Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer}
 
 \maketitle
-\thispagestyle{empty}
-\pagestyle{empty}
+%\thispagestyle{empty}
+%\pagestyle{empty}
 
-\section*{Abstract}
+\begin{abstract}
 In recent years, modular and reconfigurable manufacturing systems have gained attention for their adaptability. 
@@
 We present a straightforward optimization approach to demonstrate potential automation in this area.
+\keywords{flexible manufacturing systems, sensor placement optimization, laser tracker, particle swarm optimization}
+\end{abstract}
 \section{Introduction}
 
+The global market for flexible manufacturing systems (FMS) is experiencing significant growth, driven by the increasing demand for customized products and shorter product life cycles~\cite{westkaemperEinfuehrungOrganisationProduktion2006}.
+FMS are characterized by their ability to adapt to changing requirements, such as variations in product design, production volume, and production sequence.
+Modular reconfigurable machine tools (MRMT) are a key component of FMS, as they can modify their physical structure to meet evolving product demands~\cite{Padayachee2012}.
+
+The majority of research in the field of MRMT has concentrated on the development of architectures that adhere to open or semi-open standards, the modular components, and methods for optimizing configurations based on performance metrics~\cite{gadallaRecentAdvancesResearch2017, xuMethodDesignModular2017}.
+
+Despite the aforementioned advancements, a significant gap in the research field remains: the verification of the performance of the aforementioned proposed configurations is often limited to simulations, which are susceptible to a reality gap.
+Alternatively, manual experimental setups using external sensors are used.
+However, these manual setups are not only time-consuming but also require expert knowledge and are susceptible to errors, rendering them inadequate for rigorous verification.
+
+To address this issue, our paper introduces an automated approach to task-based sensor placement optimization for the verification of MRMT.
+For the purposes of this study, we consider a robot cell to be an ideal test case for the proposed approach, providing an upper bound for the complexity of task trajectories.
+This is based on the recognition that although the reconfiguration of robot cells is less pronounced, the task trajectories within them are typically more complex than those typically found in MRMT applications.
 \section{State of the art}
@@
 In particular we focuss on approaches that solves what~\cite{reconfigurable_production} calls reconfiguration design problem.
-This involves either combining fixed modules to form a system or directly optimizing the sytems physical design.
+This involves either combining fixed modules to form a system or directly optimizing the systems physical design. 
 Examples include robot cell optimization problems~\cite{previous_work, stiffness_placement} and robot design optimization problems~\cite{task_synthesis, ad-hoc_manipulator, multi_objective}.  
@@
         \includegraphics[width=0.8\textwidth]{figures/visibility.png}
-        \caption{Field of view (VOW) angle range and line of sight (LOS) components of marker visibility}
+        \caption{Field of view (FOV) angle range and line of sight (LOS) components of marker visibility}
         \label{fig:visibility}
@@
 where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
-Note that $p_m$ is typically dependent on time as it is attached to a moving manufacuting system such as a robot.
+Note that $p_m$ is typically dependent on time as it is attached to a moving manufacturing system such as a robot.
 The field of view objective can similarly be defined as a binary function:
@@
 The accuracy of the lasertracker is mainly dependent on the distance between the marker and the laser tracker.
-The further away the marker is the less accurate the measurment will be. However there is also a minimum measurment distance that the laser tracker can handle.
+The further away the marker is the less accurate the measurment will be. However there is also a minimum measurment distance that the laser tracker requires to find a target.
 To encode this we can define a penalty function that penalizes markers that are too far away from the laser tracker.
@@
     \end{split}
-\end{equation}I
+\end{equation}
 
@@
 Instead we can use a penalty function that penalizes any particles that violate the constraints.
-This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library. %TODO citation needed
+This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library~\cite{2020SciPy-NMeth}.
 The penalty function can be defined as:
@@
 This extension of the pybullet physics engine allows us to simulate not only the robot but also the manufacturing process.
-For the purposes of this paper the markers are implemented as a pybullet industrial endeffector obect that is attached to the robot.
+For the purposes of this paper the markers are implemented as a pybullet industrial endeffector object that is attached to the robot.
 A image of the simulation environment is shown in figure \ref{fig:simulation}.
@@
 \section{Acknowledgements}
-
+The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
+This work was supported by the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg within the InnovationsCampus Future Mobility.
+The APC was funded by Innovationscampus Mobilität der Zukunft (ICM), grant number SdManu7.
+The robot cell in the sample test case are part of the project funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) - SFB-1574 - 471687386.
 %Bibliography

### 2024-08-05 paper/paper.tex
paper to Alex

@@
 \maketitle
-%\thispagestyle{empty}
-%\pagestyle{empty}
 
@@
 Modular reconfigurable machine tools (MRMT) are a key component of FMS, as they can modify their physical structure to meet evolving product demands~\cite{Padayachee2012}.
-
 The majority of research in the field of MRMT has concentrated on the development of architectures that adhere to open or semi-open standards, the modular components, and methods for optimizing configurations based on performance metrics~\cite{gadallaRecentAdvancesResearch2017, xuMethodDesignModular2017}.
-
 Despite the aforementioned advancements, a significant gap in the research field remains: the verification of the performance of the aforementioned proposed configurations is often limited to simulations, which are susceptible to a reality gap.
@@
 However, these manual setups are not only time-consuming but also require expert knowledge and are susceptible to errors, rendering them inadequate for rigorous verification.
-
 To address this issue, our paper introduces an automated approach to task-based sensor placement optimization for the verification of MRMT.
@@
 All of these approaches focus on phyiscal capabilities such as payload, accuracy, stiffness, force, weight, and motion.
-
 To verify that the optimized system configuration achieves the desired performance, we need to measure the system's performance using external sensors,
@@
         \centering
-        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
+        \includegraphics[width=0.7\textwidth]{figures/visibility.png}
         \caption{Field of view (FOV) angle range and line of sight (LOS) components of marker visibility}
@@
 To avoid unrealistic positions, we constrain this space to keep the marker close to the manufacturing system's surface.
-This constraint can be represented cylindrically:
-\begin{equation}
-    \begin{split}
-        \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
-        0 \leq p_{p_m,z} &\leq h
-    \end{split}
-\end{equation}
-
-where $r$ is the cylinder's radius and $h$ is its height.
-Additionally, to ensure the marker remains near the surface, a penalty function can be added to the objective function to penalize markers positioned too far from the flange.\\
-\\
+This can be formulated as a cylindrical constraint.
+Additionally, to ensure the marker remains near the surface, a penalty function can be added to the objective function to penalize markers positioned too far from the flange.
 The laser tracker's pose is defined by a transformation matrix $T_{l}$, but due to its spherical measurement volume, only its position ($p_{p_l}$) needs to be specified.
@@
 \end{equation}
+where $r$ is the cylinder's radius and $h$ is its height.
 
@@
 
-\begin{table}  %TODO anpassen
+\begin{table}
+    \centering
+    \caption{Hyperparameters of the PSO algorithm}
+    \begin{minipage}{.45\linewidth}
         \centering
@@
             Number of iterations & 100 \\
+        \end{tabular}
+    \end{minipage}%
+    \hspace{0.01\linewidth}%
+    \begin{minipage}{.45\linewidth}
+        \centering
+        \begin{tabular}{c|c}
+            Hyperparameter & Value \\
+            \hline
             Inertia weight & 0.8 \\
@@
         \end{tabular}
-        \caption{Hyperparameters of the PSO algorithm}
+    \end{minipage}
+
     \label{tab:hyperparameters}
@@
 For the purposes of this paper the markers are implemented as a pybullet industrial endeffector object that is attached to the robot.
-A image of the simulation environment is shown in figure \ref{fig:simulation}.
-\begin{figure}
-        \centering
-        \includegraphics[width=\textwidth]{figures/simulation.png}
-        \caption{Simulation environment used for the optimization}
-        \label{fig:simulation}
-\end{figure}
+A image of the simulation environment is shown in Fig. \ref{fig:trajectory}.
 
@@
 The right robot services the CNC machine and the additive manufacturing system, while the left robot performs a high-precision manufacturing task.
-For this test, the left robot follows a rectangular trajectory with dimensions of 0.4 meters in length and 0.6 meters in width, as depicted in Figure \ref{fig:trajectory}.
+For this test, the left robot follows a rectangular trajectory with dimensions of 0.4 meters in length and 0.6 meters in width, as depicted in Fig. \ref{fig:trajectory}.
 The aim is now to find a laser tracker and marker placement that can track the left robot throughout its entire trajectory.
 We compare this to the performance of the initial population of marker and laser tracker placements used to start the optimization process.
-Figure \ref{fig:iterations} illustrates the performance of the algorithm over multiple iterations.
+Fig. \ref{fig:iterations} illustrates the performance of the algorithm over multiple iterations.
 The results show that the algorithm converges to a solution in roughly 40 iterations, which is relatively fast for an optimization approach.
 This speed is likely due to the existence of numerous optimal placements for the laser tracker and markers.
-The final solution is also depicted in Figure \ref{fig:trajectory}.
+The final solution is also depicted in Fig. \ref{fig:trajectory}.
 In conclusion the results show that the optimization approach is capable of finding a suitable laser tracker and marker placement for the manufacturing system.
@@
 \centering
-\includegraphics[width=\textwidth]{figures/trajectory.png}
+\includegraphics[width=0.75\textwidth]{figures/trajectory.png}
 \caption{Trajectory of the robot and final laser tracker and marker placements.}
@@
 \section{Acknowledgements}
-The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
 This work was supported by the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg within the InnovationsCampus Future Mobility.
-The APC was funded by Innovationscampus Mobilität der Zukunft (ICM), grant number SdManu7.
-The robot cell in the sample test case are part of the project funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) - SFB-1574 - 471687386.
+It was also funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) - SFB-1574 - 471687386.
 %Bibliography

### 2024-08-06 paper/paper.tex
updated trajectory.png, ready for submission

@@
 \author{Jan Baumgärtner \inst{1,2},
-Max Goebels \inst{1,2},   Alexander Puchta \inst{1,2} and Jürgen Fleischer\inst{1}% <-this % stops a space
+Max Goebels \inst{1,2,3},   Alexander Puchta \inst{1,2} and Jürgen Fleischer\inst{1}% <-this % stops a space
 }
 \institute{wbk Institute of Production Science, Karlsruhe Institute of Technology,
-76131 Karlsruhe, Germany \email{max.goebels@kit.edu}
+76131 Karlsruhe, Germany 
 \and
 These authors contributed equally to this work.
-}
+\and
+Corresponding: \email{max.goebels@kit.edu}
+}%
 
-\authorrunning{Baumgärtner et al.}
-\tocauthor{Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer}
+\authorrunning{Baumgärtner et al.}%
+\tocauthor{Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer}%
 
@@
 \begin{abstract}
-In recent years, modular and reconfigurable manufacturing systems have gained attention for their adaptability.
+In recent years, modular and reconfigurable manufacturing systems have gained attention for their fast adaptability towards product changes.
 While system design is now highly automated, designing the verification setup remains a manual design problem involving tasks like placing external sensors to measure performance.
 This paper argues that design optimization approaches for manufacturing systems can also automate the design of the verification setup.
-This is demonstrated using the placement of a laser tracker system with markers that due to their line-of-sight complexity and simultaneous measurement of multiple systems can be challenging to design.
+This is demonstrated by placing a laser tracker system with markers, which can be challenging to design due to their line-of-sight complexity and simultaneous measurement of multiple systems.
 We present a straightforward optimization approach to demonstrate potential automation in this area.
@@
 \section{Introduction}
-
 The global market for flexible manufacturing systems (FMS) is experiencing significant growth, driven by the increasing demand for customized products and shorter product life cycles~\cite{westkaemperEinfuehrungOrganisationProduktion2006}.
 FMS are characterized by their ability to adapt to changing requirements, such as variations in product design, production volume, and production sequence.
-Modular reconfigurable machine tools (MRMT) are a key component of FMS, as they can modify their physical structure to meet evolving product demands~\cite{Padayachee2012}.
+Modular reconfigurable machine tools (MRMT) are a key component of FMS, as they can modify their physical structure to meet evolving product requirements~\cite{Padayachee2012}.
 The majority of research in the field of MRMT has concentrated on the development of architectures that adhere to open or semi-open standards, the modular components, and methods for optimizing configurations based on performance metrics~\cite{gadallaRecentAdvancesResearch2017, xuMethodDesignModular2017}.
@@
 Alternatively, manual experimental setups using external sensors are used.
-However, these manual setups are not only time-consuming but also require expert knowledge and are susceptible to errors, rendering them inadequate for rigorous verification.
+However, these manual setups are not only time-consuming but also require expert knowledge.
 To address this issue, our paper introduces an automated approach to task-based sensor placement optimization for the verification of MRMT.
@@
 The goal here is to find the optimal sensor placement for accurate system performance measurement.
-A adjecent engineering discipline in which sensor placement optimization is a already well studied is structural health monitoring~\cite{shms}.
+A adjacent engineering discipline in which sensor placement optimization is a already well studied is structural health monitoring~\cite{shms}.
 Here key algorithms where already identified such as genetic algorithms, particle swarm optimization, and simulated annealing~\cite{shms}.
@@
 \section{Formulation of the optimization problem}
-To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
-Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
+To solve the joint optimization problem of marker and laser tracker pose, we first need to define the objective function.
+Our goal is to maximize the visibility of the marker to the laser tracker.
 Similar to approaches such as~\cite{ieee_sensors} we divide the visibility into two components.
@@
 The hyperparameters we need to tune are collected in table~\ref{tab:hyperparameters}.
-Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
+Note that the PSO algorithm has no direct way to encode the constraints, we defined earlier.
 Instead we can use a penalty function that penalizes any particles that violate the constraints.
-This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library~\cite{2020SciPy-NMeth}.
+This is a common approach in optimization and is also used in the implementation of the PSO algorithm in the scipy library~\cite{2020SciPy-NMeth}.
 The penalty function can be defined as:
@@
 A image of the simulation environment is shown in Fig. \ref{fig:trajectory}.
-
 \section{Sample Testcase}
@@
 We have also verified that the algorithm is capable of handling the dynamic nature of the manufacturing system, as the optimization is performed over the entire trajectory of the robot.
-
 \begin{figure}
     \centering
-\includegraphics[width=\textwidth]{figures/iterations.pdf}
+    \includegraphics[width=0.9\textwidth]{figures/trajectory.png}
+    \caption{Trajectory of the robot and final laser tracker and marker placements.}
+    \label{fig:trajectory}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\textwidth]{figures/iterations.pdf}
     \caption{Performance of the algorithm over the iterations.}
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.75\textwidth]{figures/trajectory.png}
-\caption{Trajectory of the robot and final laser tracker and marker placements.}
-\label{fig:trajectory}
-\end{figure}
-
 \section{Conclusion}
-
 In this paper, we argue that automating the verification setup in manufacturing systems is feasible using optimization methods.
